---
title: "High-Throughput Bond Length Analysis of the AMCSD"
author: "Using the Built-in crystract Batch Processor"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = TRUE, warning = TRUE)
```

## 1. Introduction

This notebook demonstrates a robust and scalable workflow for extracting and analyzing interatomic bond lengths from the entire American Mineralist Crystal Structure Database (AMCSD). This version has been updated to use the `crystract` package's high-level `analyze_cif_files` function, which provides built-in, optimized support for parallel batch processing to disk. This simplifies the code and makes the analysis more efficient and maintainable.

## 2. Setup: Loading Libraries

First, we load the necessary R packages for data manipulation, crystallographic analysis, parallel computation, and visualization.

```{r load-libraries}
# For data manipulation and speed
library(data.table)
library(stringr)
library(tidyr)

# For crystallographic analysis
library(crystract)

# For parallel processing
library(future)
library(future.apply)

# For visualization
library(ggplot2)
```

## 3. Data Preparation: One-Time Parse of the AMCSD File

The entire AMCSD dataset is provided as a single text file (`cifdata.txt`). The first step is to parse this file into a list of individual crystal structures that `crystract` can process.

**This is a time-consuming, one-time operation.** The results will be saved as an `.rds` file to avoid repeating it. If you have already run this and have the `amcsd_parsed_list.rds` file, you can skip to section 5.

```{r define-parsing-function}
#' @title Split a large multi-entry CIF file with validation
#' @description Reads, sanitizes, and splits a multi-entry CIF file.
#' @param file_path The full path to the large text file.
#' @return A named list where each element is a data.table representing one CIF file's content.
split_amcsd_file <- function(file_path) {
  if (!file.exists(file_path)) {
    stop("AMCSD data file not found at the specified path: ", file_path)
  }
  
  # Read and sanitize the file at the byte level to handle non-UTF8 characters
  con <- file(file_path, "rb")
  raw_bytes <- readBin(con, "raw", n = file.size(file_path))
  close(con)
  valid_bytes_mask <- (raw_bytes >= 32 & raw_bytes <= 126) | raw_bytes %in% as.raw(c(9, 10, 13))
  raw_bytes[!valid_bytes_mask] <- as.raw(32)
  clean_text <- rawToChar(raw_bytes, multiple = FALSE)
  
  # Split the sanitized text into individual CIF entries
  cif_entries_text <- str_split(clean_text, "(?=data_global)")[[1]]
  cif_entries_text <- cif_entries_text[str_trim(cif_entries_text) != ""]
  
  # Process each text entry into a data.table
  cif_list <- lapply(cif_entries_text, function(entry) {
    dt <- suppressWarnings(
      fread(text = entry, sep = "\n", header = FALSE, strip.white = FALSE),
      classes = "fread_incomplete_final_line"
    )
    
    # Extract the AMCSD code to use as a unique name
    amcsd_line <- dt[str_detect(V1, "_database_code_amcsd")]
    amcsd_code <- if (nrow(amcsd_line) > 0) {
      str_trim(str_replace(amcsd_line$V1, "_database_code_amcsd", ""))
    } else {
      paste0("unknown_", sample(1E6, 1))
    }
    
    list(name = amcsd_code, content = dt)
  })
  
  # Create the final named list
  cif_contents <- lapply(cif_list, `[[`, "content")
  names(cif_contents) <- sapply(cif_list, `[[`, "name")
  
  return(cif_contents)
}
```

Now, execute the parsing function. **This chunk is set to `eval=FALSE` to prevent re-running it.**

```{r execute-parsing, eval=TRUE}
# This chunk should only be run ONCE. Change eval=TRUE to run it.
amcsd_file_path <- "cifdata.txt" 
cat("Starting the parsing of the main cifdata.txt file. This may take several minutes...\n")

# Parse the file into a list of data.tables
cif_data_list <- split_amcsd_file(amcsd_file_path)

cat(sprintf("Successfully parsed %d crystal structures from the file.\n", length(cif_data_list)))

# Save the parsed list to an RDS file for quick loading in the future
saveRDS(cif_data_list, "amcsd_parsed_list.rds")
cat("Parsed data has been saved to 'amcsd_parsed_list.rds'.\n")
```

## 4. Caching Data: Saving to Individual CIF Files (Optional)

To completely avoid re-parsing `cifdata.txt` in the future, this optional step saves each crystal structure into its own `.cif` file inside the `split_cif_files/` directory.

```{r save-split-files, eval=TRUE}
# This chunk should only be run ONCE. Change eval=TRUE to run it.
output_dir <- "split_cif_files"
if (!dir.exists(output_dir)) {
  dir.create(output_dir)
}

# Load the parsed list if it's not in memory
if (!exists("cif_data_list")) {
  cif_data_list <- readRDS("amcsd_parsed_list.rds")
}

cat(sprintf("Saving %d individual CIF files to the '%s/' directory...\n", length(cif_data_list), output_dir))

# Loop through the list and write each data.table to a file
invisible(lapply(names(cif_data_list), function(name) {
  file_path <- file.path(output_dir, paste0(name, ".cif"))
  fwrite(cif_data_list[[name]], file_path, col.names = FALSE, quote = FALSE)
}))

cat("All individual CIF files have been saved.\n")
```

## 5. Loading Data for Analysis

Load the pre-parsed data from the RDS file. This is the fastest way to start the analysis on subsequent runs.

```{r load-from-rds}
# Load the pre-parsed list of CIF data
cif_data_list <- readRDS("amcsd_parsed_list.rds")
cat(sprintf("Successfully loaded %d parsed structures from 'amcsd_parsed_list.rds'.\n", length(cif_data_list)))
```

## 6. Filtering for Specific Minerals

This chunk filters the full AMCSD dataset to retain only the structures corresponding to the list of minerals provided. The subsequent analysis will only run on these selected files.

```{r filter-mineral-list}
cat("Filtering the database for the specified list of minerals...\n")

# Define the list of minerals to include in the analysis
target_minerals <- c(
  "Pyrite", "Marcasite", "Vaesite", "Cattierite", "Hauerite", "Krutovite", "krutaite", 
  "Trogtalite", "Ferroselite", "Covelline", "Arsenopyrite", "Cobaltite", "Realgar",
  "Pararealgar", "Orpiment", "Tetrahedrite", "Tennantite", "Umangite", "Calaverite",
  "Krennerite", "Sylvanite", "Montbrayite", "Melonite", "Wehrlite"
)

# Create a single regex pattern for case-insensitive matching of any mineral name
filter_pattern <- paste(target_minerals, collapse = "|")

# This function checks if a CIF's text content contains any of the target mineral names
contains_target_mineral <- function(cif_dt) {
  any(str_detect(cif_dt$V1, regex(filter_pattern, ignore_case = TRUE)))
}

# Apply the filter function across the list to get a logical vector
indices_to_keep <- sapply(cif_data_list, contains_target_mineral)

# Subset the original list to keep only the matching entries
cif_data_list <- cif_data_list[indices_to_keep]

cat(sprintf("Filtering complete. The dataset for analysis now contains %d structures.\n", length(cif_data_list)))
```

## 7. Parallel Analysis Using `crystract`'s Batch Processor

This is the most computationally intensive step. Instead of a manual loop, we now use the high-level `analyze_cif_files` function on our filtered list of minerals.

```{r run-package-batch-analysis, eval=TRUE}
# --- Set Analysis Parameters ---
num_workers <- parallel::detectCores() - 1 # Use all but one core
analysis_output_dir <- "crystract_batch_results" # Directory to save batch files
batch_processing_size <- 4000 # Number of files per batch

# --- Run the Analysis ---
cat(sprintf("Starting analysis of %d files using crystract's built-in batch processor...\n", length(cif_data_list)))
overall_time <- system.time({

  # This single function call now robustly handles parallel processing and
  # batching internally without memory overload issues.
  analyze_cif_files(
    file_paths = cif_data_list,                 # Use the pre-parsed and filtered list
    workers = num_workers,                      # Number of parallel cores
    output_dir = analysis_output_dir,           # Activate batch-to-disk mode
    batch_size = batch_processing_size,         # Files per batch
    
    # Arguments passed down to analyze_single_cif for each file
    bonding_algorithms = "minimum_distance",
    calculate_bond_angles = FALSE,
    perform_error_propagation = FALSE
  )
  
})

cat("----------------------------------\n")
cat("Package Batch Analysis Complete!\n")
cat(sprintf("Total processing time: %.2f minutes\n", overall_time["elapsed"] / 60))
cat(sprintf("Batch results have been saved in the '%s/' directory.\n", analysis_output_dir))

# The new analyze_cif_files function automatically shuts down its parallel
# workers when it completes. This call is therefore optional but is good 
# practice to include for clarity.
plan(sequential)
```

## 8. Data Aggregation and Tidying

With the analysis complete, we now use the `aggregate_batch_results` helper function from `crystract` to load and combine all the generated batch files into a single `data.table`.

```{r aggregate-data}
# The directory where the batch files were saved
analysis_output_dir <- "crystract_batch_results"

if (dir.exists(analysis_output_dir)) {
  cat(sprintf(
    "Aggregating batch results from '%s/'...\n",
    analysis_output_dir
  ))
  
  # Use the package's built-in aggregator
  all_results <- aggregate_batch_results(
    analysis_output_dir,
    cols_to_keep = c("file_name", "bonded_pairs_minimum_distance")
  )
  total_structures_processed <- nrow(all_results)
  
  # Unnest the bonded pairs data
  bonds_to_unnest <- all_results[!sapply(bonded_pairs_minimum_distance, is.null), .(file_name, bonded_pairs = bonded_pairs_minimum_distance)]
  
  if (nrow(bonds_to_unnest) > 0) {
    final_bonds_data <- unnest(bonds_to_unnest, cols = c(bonded_pairs))
    setDT(final_bonds_data)
    
    # Extract element symbols from atom labels (e.g., 'Si1' -> 'Si')
    final_bonds_data[, `:=`(
      Element1 = str_extract(Atom1, "^[A-Za-z]+"),
      Element2 = str_extract(Atom2, "^[A-Za-z]+")
    )]
    
    # Keep only needed columns and remove rows with missing data or outliers
    final_bonds_data <- final_bonds_data[, .(Element1, Element2, Distance)]
    final_bonds_data <- na.omit(final_bonds_data)
    final_bonds_data <- final_bonds_data[Distance > 0.5 &
                                           Distance < 4.5]
    
    cat(
      sprintf(
        "Aggregation complete. A total of %d unique bonds from %d structures are ready for plotting.\n",
        nrow(final_bonds_data),
        total_structures_processed
      )
    )
  } else {
    stop("Aggregation failed: No bond data found to unnest.")
  }
} else {
  stop(
    sprintf(
      "Analysis results directory '%s' not found. Please run the parallel analysis chunk first.",
      analysis_output_dir
    )
  )
}
```

## 9. Diagnostic: Identifying Failed Files

Now that we have the results, we can compare the list of successfully processed files against our original input list to identify which ones failed.

```{r identify-failed-files}
# Get the list of all file names that were initially parsed
all_input_files <- names(cif_data_list)

# Get the list of file names that were successfully processed and aggregated
processed_files <- all_results$file_name

# Use setdiff() to find the files that are in the input list but not in the results
failed_files <- setdiff(all_input_files, processed_files)

if (length(failed_files) > 0) {
  cat(sprintf(
    "Found %d files that were parsed but failed during the analysis step.\n",
    length(failed_files)
  ))
  
  # Create a data.table for better viewing
  failed_files_dt <- data.table(failed_file_name = failed_files)
  
  # Display the list in an interactive table
  DT::datatable(
    failed_files_dt,
    caption = "List of Files That Failed During Analysis",
    options = list(pageLength = 15, scrollX = TRUE),
    rownames = FALSE
  )
  
} else {
  cat("Congratulations! All parsed files were processed successfully.\n")
}
```

## 10. Visualization: Bond Length Distributions

Finally, we can visualize the results. The plot below shows the distribution of all calculated bond lengths for each chemical element, ordered by atomic number.

```{r create-plot, fig.width=16, fig.height=9}
if (exists("final_bonds_data") && nrow(final_bonds_data) > 0) {
  # Reshape the data into a long format for ggplot
  plot_data_long <- rbind(
    final_bonds_data[, .(Element = Element1, Distance)], 
    final_bonds_data[, .(Element = Element2, Distance)]
  )
  
  # Define atomic numbers to order the elements logically
  atomic_numbers <- c(H=1, He=2, Li=3, Be=4, B=5, C=6, N=7, O=8, F=9, Ne=10, Na=11, Mg=12, Al=13, Si=14, P=15, S=16, Cl=17, Ar=18, K=19, Ca=20, Sc=21, Ti=22, V=23, Cr=24, Mn=25, Fe=26, Co=27, Ni=28, Cu=29, Zn=30, Ga=31, Ge=32, As=33, Se=34, Br=35, Kr=36, Rb=37, Sr=38, Y=39, Zr=40, Nb=41, Mo=42, Tc=43, Ru=44, Rh=45, Pd=46, Ag=47, Cd=48, In=49, Sn=50, Sb=51, Te=52, I=53, Xe=54, Cs=55, Ba=56, La=57, Ce=58, Pr=59, Nd=60, Pm=61, Sm=62, Eu=63, Gd=64, Tb=65, Dy=66, Ho=67, Er=68, Tm=69, Yb=70, Lu=71, Hf=72, Ta=73, W=74, Re=75, Os=76, Ir=77, Pt=78, Au=79, Hg=80, Tl=81, Pb=82, Bi=83, Po=84, At=85, Rn=86, Fr=87, Ra=88, Ac=89, Th=90, Pa=91, U=92, Np=93, Pu=94)
  
  # Filter data and set the factor levels for ordering
  plot_data_long <- plot_data_long[Element %in% names(atomic_numbers)]
  plot_data_long$Element <- factor(plot_data_long$Element, levels = names(sort(atomic_numbers)))
  
  # Generate the boxplot
  bond_length_plot <- ggplot(plot_data_long, aes(x = Element, y = Distance)) +
    geom_boxplot(outlier.shape = NA, fill = "#4682B4", color = "black") +
    coord_cartesian(ylim = c(0.5, 4.0)) + # Zoom in on the most common range
    labs(
      title = "Distribution of Interatomic Bond Lengths Across the Periodic Table",
      subtitle = paste("Based on", format(total_structures_processed, big.mark = ","), "structures from the AMCSD"),
      x = "Element",
      y = "Bond Length (Ångströms)"
    ) +
    theme_minimal(base_size = 14) +
    theme(
      axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 10),
      panel.grid.major.x = element_blank(),
      plot.title = element_text(face = "bold", size = 20),
      plot.subtitle = element_text(size = 16)
    )
  
  print(bond_length_plot)
} else {
  cat("No data available to create the bond length plot.\n")
}
```

### 10.2 Heatmap

```{r}
`

```

## 11. Conclusion

This workflow successfully demonstrates a scalable and efficient method for analyzing a massive crystallographic database. By leveraging `crystract`'s built-in, parallelized batch processing, we greatly simplified the analysis code while improving its robustness and memory efficiency. The final result is a comprehensive visualization of chemical bonding trends across the periodic table, derived from over 21,000 crystal structures.
