---
title: "Project Plan"
author: "Don Ngo"
date: "`r Sys.Date()`"
output: pdf_document
---

#Packages (to be adjusted)

```{r}
library(cry)
library(data.table)
library(readtext)
library(stringr)
```

# General Objectives

-   Create a private GitHub repository.
-   Extract information from crystallography files.
-   Identify necessary information for cluster analyses.
-   Build functions.
-   Build a package.

# Specific Objectives and Tasks

## Task 1: Create a Private GitHub Repository

**Action:** Set up a private repository on GitHub.\
**Requirement:** Each code segment added should be accompanied by an explanation.

## Task 2: Bind CIF Files into One Table

**Steps:**

1.  **Download CIF files from the provided Google Drive link.**

2.  **Set the working directory:**

```{r}
setwd("~/repos/ml-crystals/")
```

3.  **List CIF files**

```{r}
#cif_list <- list.files(pattern="\\.cif$", full.names = TRUE, ignore.case = TRUE)
cif_list <- Sys.glob(paths = "data/*.cif")
cif_tab <- lapply(cif_list, readLines , warn =  FALSE)
```

##Task 3: Extracting Data

```{r}
# Extracting the chemical formula
extract_chemical_formula <- function(cif_content) {
  formula_lines <- grep("_chemical_formula_sum", cif_content, value = TRUE)
  if (length(formula_lines) > 0) {
    chemical_formula <- gsub("_chemical_formula_sum", "", formula_lines)
    chemical_formula <- gsub("'", "", chemical_formula)
    chemical_formula <- trimws(chemical_formula)
    return(chemical_formula)
  } else {
    return(NA)
  }
}
# Extracting the structure type
extract_structure_type <- function(cif_content) {
  structure_lines <- grep("_chemical_name_structure_type", cif_content, value = TRUE)
  if (length(structure_lines) > 0) {
    structure_type <- gsub("_chemical_name_structure_type", "", structure_lines)
    structure_type <- gsub("'", "", structure_type)
    structure_type <- trimws(structure_type)
    return(structure_type)
  } else {
    return(NA)
  }
}
# Extracting the space group
extract_space_group_name <- function(cif_content) {
  space_group_lines <- grep("_space_group_name_H-M_alt", cif_content, value = TRUE)
  if (length(space_group_lines) > 0) {
    space_group <- gsub("_space_group_name_H-M_alt", "", space_group_lines)
    space_group <- gsub("'", "", space_group)
    space_group <- trimws(space_group)
    return(space_group)
  } else {
    return(NA)
  }
}

extract_space_group_number <- function(cif_content) {
  space_group_number_lines <- grep("_space_group_IT_number", cif_content, value = TRUE)
  if (length(space_group_number_lines) > 0) {
    space_group_number <- gsub("_space_group_IT_number", "", space_group_number_lines)
    space_group_number <- trimws(space_group_number)
    return(space_group_number)
  } else {
    return(NA)
  }
}

# Extracting reference info
# For each _citation_ tag, extract the authors, title, journal, year, volume, and pages
# _citation_author_name, _citation_title, _citation_journal_full, _citation_year, _citation_journal_volume, _citation_page_first, _citation_page_last
# Return a list of reference information
# Multiple references are possible
# references can be found next to the _citation_ tag or under the _citation_ tag until the next tag


# Function to extract unit cell length and associated errors
extract_unit_cell_metrics <- function(cif_content) {
  extract_value_with_error <- function(line) {
    if (grepl("\\(", line)) {
      value <- sub("\\(.*\\)", "", line)
      error <- sub(".*\\((.*)\\).*", "\\1", line)
      return(list(value = as.numeric(value), error = as.numeric(error)))
    } else {
      return(list(value = as.numeric(line), error = NA))
    }
  }

  a_line <- grep("_cell_length_a", cif_content, value = TRUE)
  b_line <- grep("_cell_length_b", cif_content, value = TRUE)
  c_line <- grep("_cell_length_c", cif_content, value = TRUE)

  a <- extract_value_with_error(gsub("_cell_length_a", "", a_line))
  b <- extract_value_with_error(gsub("_cell_length_b", "", b_line))
  c <- extract_value_with_error(gsub("_cell_length_c", "", c_line))

  return(list(
    a = a$value, a_error = a$error,
    b = b$value, b_error = b$error,
    c = c$value, c_error = c$error
  ))
}
```

Extracting from all the .cifs:

```{r}
# Initialize an empty list to store the results
results <- list()

for (cif_file in cif_list) {
  # Read the CIF file
  cif_content <- readLines(cif_file, warn = FALSE)

  # Extract the compound name from the file name
  compound_name <- tools::file_path_sans_ext(basename(cif_file))

  # Extract the relevant data
  chemical_formula <- extract_chemical_formula(cif_content)
  structure_type <- extract_structure_type(cif_content)
  space_group_name <- extract_space_group_name(cif_content)
  space_group_number <- extract_space_group_number(cif_content)
  #reference_info <- extract_reference_info(cif_content)
  unit_cell_metrics <- extract_unit_cell_metrics(cif_content)

  # Combine the data into a single list
  result <- list(
    compound_name = compound_name,
    chemical_formula = chemical_formula,
    structure_type = structure_type,
    space_group_name = space_group_name,
    space_group_number = space_group_number,
    #reference_info = reference_info,
    unit_cell_metrics = unit_cell_metrics
  )

  # Append the result to the results list
  results <- append(results, list(result))
}

# Convert the results list to a data frame
results_df <- do.call(rbind, lapply(results, as.data.frame))
```
