---
title: "Project Plan"
author: "Don Ngo"
date: "`r Sys.Date()`"
output: pdf_document
---

#Packages (to be adjusted)

```{r}
library(cry)
library(data.table)
library(readtext)
library(stringr)
```

# General Objectives

-   Create a private GitHub repository.
-   Extract information from crystallography files.
-   Identify necessary information for cluster analyses.
-   Build functions.
-   Build a package.

# Specific Objectives and Tasks

## Task 1: Create a Private GitHub Repository

**Action:** Set up a private repository on GitHub.\
**Requirement:** Each code segment added should be accompanied by an explanation.

## Task 2: Bind CIF Files into One Table

**Steps:**

1.  **Download CIF files from the provided Google Drive link.**

2.  **Set the working directory:**

```{r}
setwd("~/repos/ml-crystals/")
```

3.  **List CIF files**

```{r}
cif_list <- Sys.glob(paths = "data/*.cif")
cif_tab <- lapply(cif_list, readLines , warn =  FALSE)
```

##Task 3 and 4: Extracting Data

```{r}
# Extracting the chemical formula
extract_chemical_formula <- function(cif_content) {
  formula_lines <- grep("_chemical_formula_sum", cif_content, value = TRUE)
  if (length(formula_lines) > 0) {
    chemical_formula <- gsub("_chemical_formula_sum", "", formula_lines)
    chemical_formula <- gsub("'", "", chemical_formula)
    chemical_formula <- trimws(chemical_formula)
    return(chemical_formula)
  } else {
    return(NA)
  }
}
# Extracting the structure type
extract_structure_type <- function(cif_content) {
  structure_lines <- grep("_chemical_name_structure_type", cif_content, value = TRUE)
  if (length(structure_lines) > 0) {
    structure_type <- gsub("_chemical_name_structure_type", "", structure_lines)
    structure_type <- gsub("'", "", structure_type)
    structure_type <- trimws(structure_type)
    return(structure_type)
  } else {
    return(NA)
  }
}
# Extracting the space group
extract_space_group_name <- function(cif_content) {
  space_group_lines <- grep("_space_group_name_H-M_alt", cif_content, value = TRUE)
  if (length(space_group_lines) > 0) {
    space_group <- gsub("_space_group_name_H-M_alt", "", space_group_lines)
    space_group <- gsub("'", "", space_group)
    space_group <- trimws(space_group)
    return(space_group)
  } else {
    return(NA)
  }
}

extract_space_group_number <- function(cif_content) {
  space_group_number_lines <- grep("_space_group_IT_number", cif_content, value = TRUE)
  if (length(space_group_number_lines) > 0) {
    space_group_number <- gsub("_space_group_IT_number", "", space_group_number_lines)
    space_group_number <- trimws(space_group_number)
    return(space_group_number)
  } else {
    return(NA)
  }
}

# Function to extract unit cell length and associated errors
extract_unit_cell_metrics <- function(cif_content) {
  extract_value_with_error <- function(line) {
    if (grepl("\\(", line)) {
      value <- sub("\\(.*\\)", "", line)
      error <- sub(".*\\((.*)\\).*", "\\1", line)
      return(list(value = as.numeric(value), error = as.numeric(error)))
    } else {
      return(list(value = as.numeric(line), error = NA))
    }
  }

  a_line <- grep("_cell_length_a", cif_content, value = TRUE)
  b_line <- grep("_cell_length_b", cif_content, value = TRUE)
  c_line <- grep("_cell_length_c", cif_content, value = TRUE)

  a <- extract_value_with_error(gsub("_cell_length_a", "", a_line))
  b <- extract_value_with_error(gsub("_cell_length_b", "", b_line))
  c <- extract_value_with_error(gsub("_cell_length_c", "", c_line))

  return(list(
    a = a$value, a_error = a$error,
    b = b$value, b_error = b$error,
    c = c$value, c_error = c$error
  ))
}

```

## Task 5: extracting atomic coordinates and symmetry operations and applying symmetry operations to the atomic coordinates

``` {r}
# Extracting atomic coordinates
extract_atomic_coordinates <- function(cif_content) {
  # Initialize lists to store atom properties
  labels <- vector()
  fractional_x <- vector()
  fractional_y <- vector()
  fractional_z <- vector()
  x_errors <- vector()
  y_errors <- vector()
  z_errors <- vector()

  # Find the start and end of the atom section
  atom_start <- grep("_atom_site_occupancy", cif_content)
  atom_end <- grep("^loop_|^#End", cif_content[atom_start:length(cif_content)]) + atom_start - 1

  # If atom_end has multiple matches, take the first match after atom_start
  if(length(atom_end) > 1) {
    atom_end <- atom_end[1]
  }

  # Iterate over each line containing atomic coordinates
  for (line in cif_content[(atom_start + 1):(atom_end - 1)]) {
    # Extract atom properties from the line
    properties <- strsplit(line, "\\s+")[[1]]

    # Extract and store relevant information
    labels <- c(labels, properties[1])

    # Extract fractional coordinates and errors separately
    fractional_x_val <- gsub("\\(.*\\)", "", properties[5])
    fractional_y_val <- gsub("\\(.*\\)", "", properties[6])
    fractional_z_val <- gsub("\\(.*\\)", "", properties[7])

    x_error <- ifelse(grepl("\\(", properties[5]), gsub(".*\\((.*)\\).*", "\\1", properties[5]), NA)
    y_error <- ifelse(grepl("\\(", properties[6]), gsub(".*\\((.*)\\).*", "\\1", properties[6]), NA)
    z_error <- ifelse(grepl("\\(", properties[7]), gsub(".*\\((.*)\\).*", "\\1", properties[7]), NA)

    fractional_x <- c(fractional_x, fractional_x_val)
    fractional_y <- c(fractional_y, fractional_y_val)
    fractional_z <- c(fractional_z, fractional_z_val)
    x_errors <- c(x_errors, x_error)
    y_errors <- c(y_errors, y_error)
    z_errors <- c(z_errors, z_error)
  }

  # Create a data frame to store the atomic coordinates
  atomic_coordinates <- data.frame(
    Label = labels,
    x_a = fractional_x,
    y_b = fractional_y,
    z_c = fractional_z,
    x_error = x_errors,
    y_error = y_errors,
    z_error = z_errors
  )
  return(atomic_coordinates)
}

# Extracting symmetry operations
extract_symmetry_operations <- function(cif_content) {
  # Find the start and end of the symmetry operations section
  symmetry_start <- grep("_space_group_symop_operation_xyz", cif_content)
  symmetry_end <- grep("^loop_", cif_content[symmetry_start:length(cif_content)]) + symmetry_start - 1

  # Extract the symmetry operations
  symmetry_operations <- cif_content[(symmetry_start + 1):(symmetry_end - 1)]
  return(symmetry_operations)
}

#test the symmetry operations
cif_content <- readLines("data/ICSD138351.cif")
symmetry_operations <- extract_symmetry_operations(cif_content)
print(symmetry_operations)
```

## Task 6: Calculating distances between atoms

```{r}
# Insert logic here
```

# Extracting from all the .cifs:

```{r}
# Initialize an empty list to store the results
results <- list()

for (cif_file in cif_list) {
  # Read the CIF file
  cif_content <- readLines(cif_file, warn = FALSE)

  # Extract the compound name from the file name
  compound_name <- tools::file_path_sans_ext(basename(cif_file))

  # Extract the relevant data
  chemical_formula <- extract_chemical_formula(cif_content)
  structure_type <- extract_structure_type(cif_content)
  space_group_name <- extract_space_group_name(cif_content)
  space_group_number <- extract_space_group_number(cif_content)
  unit_cell_metrics <- extract_unit_cell_metrics(cif_content)
  atomic_coordinates <- extract_atomic_coordinates(cif_content)
  # Combine the data into a single list
  result <- list(
    compound_name = compound_name,
    chemical_formula = chemical_formula,
    structure_type = structure_type,
    space_group_name = space_group_name,
    space_group_number = space_group_number,
    unit_cell_metrics = unit_cell_metrics,
    atomic_coordinates = atomic_coordinates
  )

  # Append the result to the results lista
  results <- append(results, list(result))
}

# Convert the results list to a data frame
results_df <- do.call(rbind, lapply(results, as.data.frame))
print(results_df)
```
