---
title: "Developing a Machine Learning Pipeline for Crystallography Data"
author: "Don Ngo"
date: "`r Sys.Date()`"
output: pdf_document
---

# General Objectives

-   Create a private GitHub repository.
-   Extract information from crystallography files.
-   Identify necessary information for cluster analyses.
-   Build functions.
-   Build a package.

## Functions

### Set the working directory

```{r}
setwd("C:/repos/ml-crystals")
```

### List CIF files

```{r}
cif_list <- Sys.glob(paths = "data/*.cif")
cif_tab <- lapply(cif_list, readLines , warn =  FALSE)
```

### Packages (to be adjusted)

```{r}
library(data.table)
library(stringr)
library(dplyr)
```

### Bind CIF Files into One Table

```{r}
# List CIF files
cif_list <- Sys.glob(paths = "data/*.cif")
cif_tab <- lapply(cif_list, fread, sep = "\n", header = FALSE)
```

### Extracting Data

```{r}
# Function to extract a specific value based on a pattern from CIF content
extract_value <- function(cif_content, pattern, remove_pattern = TRUE) {
  # Find lines that match the given pattern
  lines <- cif_content[V1 %like% pattern]
  if (nrow(lines) > 0) {
    value <- lines$V1
    if (remove_pattern) {
      value <- gsub(pattern, "", value)
    }
    # Clean the value by removing quotes and trimming whitespace
    value <- gsub("'", "", value)
    value <- trimws(value)
    return(value)
  } else {
    return(NA)
  }
}
```

### Specific extraction functions for different properties

```{r}
extract_database_code <- function (cif_content) {
  extract_value(cif_content, "_database_code_")
}

extract_chemical_formula <- function(cif_content) {
  extract_value(cif_content, "_chemical_formula_sum")
}

extract_structure_type <- function(cif_content) {
  extract_value(cif_content, "_chemical_name_structure_type")
}

extract_space_group_name <- function(cif_content) {
  extract_value(cif_content, "_space_group_name_H-M_alt")
}

extract_space_group_number <- function(cif_content) {
  extract_value(cif_content, "_space_group_IT_number")
}
```

### Function to extract unit cell metrics

```{r}
extract_unit_cell_metrics <- function(cif_content) {
  cell_parameters <- c(
    "_cell_length_a",
    "_cell_length_b",
    "_cell_length_c",
    "_cell_angle_alpha",
    "_cell_angle_beta",
    "_cell_angle_gamma"
  )

  # Extract and clean values for each cell parameter
  values <- sapply(cell_parameters, function(param) {
    line <- cif_content[V1 %like% param]$V1
    if (length(line) > 0) {
      value <- gsub(".*\\s+([0-9\\.]+).*", "\\1", line)
      return(as.numeric(value))
    } else {
      return(NA)
    }
  })

  # Create a data table with the extracted values
  unit_cell_metrics <- as.data.table(t(values))
  setnames(unit_cell_metrics, cell_parameters)
  
  return(unit_cell_metrics)
}
```

### Extracting atomic coordinates

```{r}
extract_atomic_coordinates <- function(cif_content) {
  # Initialize lists to store atom properties
  labels <- vector()
  fractional_x <- vector()
  fractional_y <- vector()
  fractional_z <- vector()
  x_errors <- vector()
  y_errors <- vector()
  z_errors <- vector()

  # Helper function to calculate the scaled error
  scale_error <- function(value, error) {
    if (is.na(error)) return(NA)
    # Count the number of decimal places in the value
    decimal_places <- nchar(strsplit(as.character(value), "\\.")[[1]][2])
    scaled_error <- as.numeric(paste0("0.", strrep("0", decimal_places - nchar(error)), error))
    return(scaled_error)
  }

  # Find the start and end of the atom section
  atom_start <- grep("_atom_site_occupancy", cif_content$V1)
  atom_end <- grep("^loop_|^#End", cif_content$V1[atom_start:length(cif_content$V1)]) + atom_start - 1

  # If atom_end has multiple matches, take the first match after atom_start
  if (length(atom_end) > 1) {
    atom_end <- atom_end[1]
  }

  # Iterate over each line containing atomic coordinates
  for (line in cif_content$V1[(atom_start + 1):(atom_end - 1)]) {
    # Extract atom properties from the line
    properties <- strsplit(line, "\\s+")[[1]]

    # Extract and store relevant information
    labels <- c(labels, as.character(properties[1]))

    # Extract fractional coordinates and errors separately
    fractional_x_val <- gsub("\\(.*\\)", "", properties[5])
    fractional_y_val <- gsub("\\(.*\\)", "", properties[6])
    fractional_z_val <- gsub("\\(.*\\)", "", properties[7])

    x_error <- ifelse(grepl("\\(", properties[5]), gsub(".*\\((.*)\\).*", "\\1", properties[5]), NA)
    y_error <- ifelse(grepl("\\(", properties[6]), gsub(".*\\((.*)\\).*", "\\1", properties[6]), NA)
    z_error <- ifelse(grepl("\\(", properties[7]), gsub(".*\\((.*)\\).*", "\\1", properties[7]), NA)

    # Adjust the errors to match the precision of the coordinates
    fractional_x <- c(fractional_x, fractional_x_val)
    fractional_y <- c(fractional_y, fractional_y_val)
    fractional_z <- c(fractional_z, fractional_z_val)
    
    x_errors <- c(x_errors, scale_error(fractional_x_val, x_error))
    y_errors <- c(y_errors, scale_error(fractional_y_val, y_error))
    z_errors <- c(z_errors, scale_error(fractional_z_val, z_error))
  }

  # Create a data table to store the atomic coordinates
  atomic_coordinates <- data.table(
    Label = labels,
    x_a = as.numeric(fractional_x),
    y_b = as.numeric(fractional_y),
    z_c = as.numeric(fractional_z),
    x_error = as.numeric(x_errors),
    y_error = as.numeric(y_errors),
    z_error = as.numeric(z_errors)
  )
  
  return(atomic_coordinates)
}
```

### Function to extract symmetry operations from CIF content

```{r}
extract_symmetry_operations <- function(cif_content) {
  # Find the start and end of the symmetry operations section
  symmetry_start <- grep("_space_group_symop_operation_xyz", cif_content$V1)
  symmetry_end <- grep("^loop_", cif_content$V1[symmetry_start:length(cif_content$V1)]) + symmetry_start - 1

  if (length(symmetry_end) > 1) {
    symmetry_end <- symmetry_end[1]
  }

  # Extract lines containing symmetry operations
  symmetry_lines <- cif_content$V1[(symmetry_start + 1):(symmetry_end - 1)]
  # Split each line into components
  symmetry_operations <- str_split_fixed(symmetry_lines, ",\\s*", n = 3)

  # Create a data table with the extracted symmetry operations
  symmetry_operations <- data.table(
    x = gsub("^[0-9]+\\s+'|'", "", symmetry_operations[, 1]),
    y = gsub("^[0-9]+\\s+'|'", "", symmetry_operations[, 2]),
    z = gsub("^[0-9]+\\s+'|'", "", symmetry_operations[, 3])
  )

  return(symmetry_operations)
}
```

### Function to apply symmetry operations to atomic coordinates

```{r}
apply_symmetry_operations <- function(atomic_coordinates, symmetry_operations) {
  # Helper function to apply a single operation
  apply_operation <- function(operation, x, y, z) {
    operation <- gsub("x", sprintf("(%f)", x), operation)
    operation <- gsub("y", sprintf("(%f)", y), operation)
    operation <- gsub("z", sprintf("(%f)", z), operation)
    eval(parse(text = operation))
  }

  # Expand coordinates by applying each symmetry operation
  expand_coordinates <- function(row) {
    x <- row$x_a
    y <- row$y_b
    z <- row$z_c
    new_coords <- rbindlist(lapply(1:nrow(symmetry_operations), function(i) {
      new_x <- apply_operation(symmetry_operations[i, x], x, y, z)
      new_y <- apply_operation(symmetry_operations[i, y], x, y, z)
      new_z <- apply_operation(symmetry_operations[i, z], x, y, z)

      data.table(
        Label = paste(row$Label, i, sep = "_"), # Apply unique label
        #Regular label
        #Label = row$Label,
        x_a = ifelse(new_x > 1, new_x - 1, ifelse(new_x < 0, new_x + 1, new_x)),
        y_b = ifelse(new_y > 1, new_y - 1, ifelse(new_y < 0, new_y + 1, new_y)),
        z_c = ifelse(new_z > 1, new_z - 1, ifelse(new_z < 0, new_z + 1, new_z))
      )
    }))
    return(new_coords)
  }

  # Apply symmetry operations to all atomic coordinates
  transformed_coords <- rbindlist(lapply(1:nrow(atomic_coordinates), function(i) {
    expand_coordinates(atomic_coordinates[i])
  }))

  # Remove duplicates
  transformed_coords <- transformed_coords %>%
    distinct(x_a, y_b, z_c, .keep_all = TRUE)

  return(transformed_coords)
}
```

### Function to expand transformed coordinates in all neighboring unit cells
```{r}
expand_transformed_coords <- function(transformed_coords) {
  # Define the number of unit cells to expand in each direction
  n_cells <- 1

  # Generate a grid of unit cell indices to expand
  cell_indices <- as.data.table(expand.grid(
    x = -n_cells:n_cells,
    y = -n_cells:n_cells,
    z = -n_cells:n_cells
  ))

  # Expand the transformed coordinates
  expanded_coords <- rbindlist(lapply(1:nrow(cell_indices), function(i) {
    cell_shift <- cell_indices[i]
    new_coords <- transformed_coords[, .(
      Label = paste(Label, paste(cell_shift$x, cell_shift$y, cell_shift$z, sep = "_"), sep = "_"),
      #Regular label
      #Label = Label,
      x_a = x_a + cell_shift$x,
      y_b = y_b + cell_shift$y,
      z_c = z_c + cell_shift$z
    )]
    return(new_coords)
  }))

  # Remove duplicates
  expanded_coords <- expanded_coords %>%
    distinct(x_a, y_b, z_c, .keep_all = TRUE)

  return(expanded_coords)
}
```
### Function to calculate distances between atoms using real-space metric tensor and vectorization

```{r}
calculate_distances <- function(atomic_coordinates, expanded_coords, unit_cell_metrics) {
  # Extract unit cell parameters
  a <- unit_cell_metrics$`_cell_length_a`
  b <- unit_cell_metrics$`_cell_length_b`
  c <- unit_cell_metrics$`_cell_length_c`
  alpha <- unit_cell_metrics$`_cell_angle_alpha` * pi / 180  # Convert to radians
  beta <- unit_cell_metrics$`_cell_angle_beta` * pi / 180    # Convert to radians
  gamma <- unit_cell_metrics$`_cell_angle_gamma` * pi / 180  # Convert to radians

  # Extract coordinates as matrices
  coords_atomic <- as.matrix(atomic_coordinates[, .(x_a, y_b, z_c)])
  coords_expanded <- as.matrix(expanded_coords[, .(x_a, y_b, z_c)])
  labels_atomic <- atomic_coordinates$Label
  labels_expanded <- expanded_coords$Label

  # Calculate pairwise differences in coordinates between atomic and transformed coordinates
  delta_x <- outer(coords_atomic[, 1], coords_expanded[, 1], "-")
  delta_y <- outer(coords_atomic[, 2], coords_expanded[, 2], "-")
  delta_z <- outer(coords_atomic[, 3], coords_expanded[, 3], "-")

  # Cosine terms for angles
  cos_alpha <- cos(alpha)
  cos_beta <- cos(beta)
  cos_gamma <- cos(gamma)
  
  # Set near-zero cosine values to exactly 0
  if (abs(cos_alpha) < 1e-10) cos_alpha <- 0
  if (abs(cos_beta) < 1e-10) cos_beta <- 0
  if (abs(cos_gamma) < 1e-10) cos_gamma <- 0

  # Calculate the squared distance using the metric tensor
  r2 <- ((a^2) * (delta_x^2)) + ((b^2) * (delta_y^2)) + ((c^2) * (delta_z^2)) +
        (2 * b * c * cos_alpha * delta_y * delta_z) +
        (2 * c * a * cos_beta * delta_z * delta_x) +
        (2 * a * b * cos_gamma * delta_x * delta_y)

  # Calculate the distances
  r <- sqrt(r2)

  # Create a distance matrix with atom labels
  atom_pairs <- expand.grid(Atom1 = labels_atomic, Atom2 = labels_expanded, KEEP.OUT.ATTRS = TRUE)

  distances <- data.table(
    Atom1 = atom_pairs$Atom1,
    Atom2 = atom_pairs$Atom2,
    Distance = as.vector(r),
    Alpha = as.vector(alpha),
    DeltaX = as.vector(delta_x),
    DeltaY = as.vector(delta_y),
    DeltaZ = as.vector(delta_z),
    CosAlpha = as.vector(cos_alpha),
    CosBeta = as.vector(cos_beta),
    CosGamma = as.vector(cos_gamma)
  )

  distances <- distances[Distance > 0]
  # Filter out distances above 10 Angstrom if necessary
  # distances <- distances[Distance > 0.01 & Distance < 10]

  return(distances)
}

```

### Function to apply Minimum Distance Method

```{r}
minimum_distance <- function(distances, delta = 0.1) {
  # Calculate dmin for each Atom1
  dmin <- distances[, .(dmin = min(Distance)), by = .(Atom1)]
  
  # Calculate dcut for each Atom1
  dmin[, dcut := (1 + delta) * dmin, print(Atom1)]
  
  # Determine bonded pairs based on dcut
  bonded_pairs <- distances[dmin, on = .(Atom1), allow.cartesian = TRUE][Distance <= dcut, .(Atom1, Atom2, Distance, dcut, dmin)]
  
  # Remove duplicates
  # unique_bonds <- bonded_pairs %>%
  #   distinct(Distance, .keep_all = TRUE)
  
  return(bonded_pairs)
  
}
```

### Function to apply Brunner's Largest Reciprocal Gap Method

```{r}
brunner <- function(distances, delta = 0.0001) {
  # Initialize an empty list to store the bonds
  bonds <- list()
  
  # Iterate over each unique central atom in the distances table
  unique_atoms <- unique(distances$Atom1)
  for (atom in unique_atoms) {
    # Filter distances for the current Atom1 and order by distance
    atom_distances <- distances[Atom1 == atom][order(Distance)]
    
    # Initialize the largest reciprocal gap and the corresponding index
    largest_gap <- -Inf
    j_max <- NA
    
    # Calculate the largest reciprocal gap
    for (j in 1:(nrow(atom_distances) - 1)) {
      reciprocal_gap <- 1 / atom_distances$Distance[j] - 1 / atom_distances$Distance[j + 1]
      if (reciprocal_gap > largest_gap) {
        largest_gap <- reciprocal_gap
        j_max <- j
      }
    }
    
    # Determine the distance cut-off
    d_cut <- atom_distances$Distance[j_max] + delta
    
    # Select the bonds within the distance cut-off
    bonded_atoms <- atom_distances[Distance <= d_cut]
    
    # Store the bonds for the current atom
    bonds[[atom]] <- bonded_atoms
  }
  
  # Combine all bonds into a single data.table
  bonds <- rbindlist(bonds, fill = TRUE)
  return(bonds)
}
```

### Function to apply Hoppe's Method of Effective Coordination Numbers

```{r}
hoppe <- function(distances, delta = 0.5, tolerance = 0.001) {
  # Initialize an empty list to store the bonded pairs
  bonded_pairs <- list()
  
  # Iterate over each unique Atom1 in the distances table
  unique_atoms <- unique(distances$Atom1)
  for (atom in unique_atoms) {
    # Filter distances for the current Atom1
    atom_distances <- distances[Atom1 == atom]

    # Step 1: Calculate dmin for the current atom
    dmin <- min(atom_distances$Distance)
    
    # Step 2: Calculate initial davg (0davg)
    davg <- sum(atom_distances$Distance * exp(1 - (atom_distances$Distance / dmin)^6)) / 
            sum(exp(1 - (atom_distances$Distance / dmin)^6))
    
    # Step 3: Iteratively calculate ndavg until convergence
    ndavg <- davg
    while (TRUE) {
      prev_davg <- davg
      davg <- sum(atom_distances$Distance * exp(1 - (atom_distances$Distance / prev_davg)^6)) / 
              sum(exp(1 - (atom_distances$Distance / prev_davg)^6))
      if (abs(davg - prev_davg) <= tolerance) break
    }
    
    # Step 4: Determine bonded pairs and their bond strength based on the tolerance parameter δ
    atom_distances[, BondStrength := exp(1 - (Distance / davg)^6)]
    bonded <- atom_distances[BondStrength >= delta, .(Atom1, Atom2, Distance)]
    bonded_pairs[[atom]] <- bonded
  }
  
  # Combine all bonded pairs into a single data.table
  bonded_pairs <- rbindlist(bonded_pairs, fill = TRUE)
  return(bonded_pairs)
}
```

### Calculate Bond Angles

``` {r}
# # Function to calculate bond angles with debugging statements
# calculate_angles <- function(bonded_pairs, distances) {
#   # Convert relevant columns to character to avoid factor level issues
#   bonded_pairs[, Atom1 := as.character(Atom1)]
#   bonded_pairs[, Atom2 := as.character(Atom2)]
#   
#   distances[, Atom1 := as.character(Atom1)]
#   distances[, Atom2 := as.character(Atom2)]
#   
#   # Print unique atoms for debugging
#   print("Unique Atoms in Bonded Pairs:")
#   print(unique(c(bonded_pairs$Atom1, bonded_pairs$Atom2)))
#   
#   print("Unique Atoms in Distances:")
#   print(unique(c(distances$Atom1, distances$Atom2)))
#   
#   # Merge bonded pairs with distances to get the distances for each bond
#   bond_distances <- merge(bonded_pairs, distances, by = c("Atom1", "Atom2"), all.x = TRUE)
#   
#   print("Bond Distances:")
#   print(bond_distances)
#   
#   # Prepare an empty data.table to store angles
#   angles <- data.table(Atom1 = character(), Atom2 = character(), Atom3 = character(), Angle = numeric())
#   
#   # Loop through each unique central atom
#   unique_atoms <- unique(bonded_pairs$Atom1)
#   
#   print("Unique Atoms:")
#   print(unique_atoms)
#   
#   for (atom in unique_atoms) {
#     # Filter bonds where the atom is the central atom
#     central_bonds <- bond_distances[Atom1 == atom]
#     
#     print(paste("Central Atom:", atom))
#     print("Central Bonds:")
#     print(central_bonds)
#     
#     # Loop through all pairs of bonds involving the central atom
#     for (i in 1:(nrow(central_bonds) - 1)) {
#       for (j in (i + 1):nrow(central_bonds)) {
#         # Bond 1
#         atom2 <- central_bonds$Atom2[i]
#         distance1 <- central_bonds$Distance[i]
#         
#         # Bond 2
#         atom3 <- central_bonds$Atom2[j]
#         distance2 <- central_bonds$Distance[j]
#         
#         # Bond 3 (distance between atom2 and atom3)
#         print(paste("Looking for distance between:", atom2, "and", atom3))
#         bond_2_3 <- distances[Atom1 == atom2 & Atom2 == atom3 | Atom1 == atom3 & Atom2 == atom2]
#         
#         print("Bond 2-3:")
#         print(bond_2_3)
#         
#         if (nrow(bond_2_3) > 0) {
#           distance3 <- bond_2_3$Distance[1] # There should be only one distance in this case
#           
#           # Calculate the angle using cosine rule
#           cos_gamma <- (distance1^2 + distance2^2 - distance3^2) / (2 * distance1 * distance2)
#           cos_gamma <- pmin(pmax(cos_gamma, -1), 1) # Ensure cos_gamma is within valid range
#           gamma <- acos(cos_gamma) * 180 / pi # Convert to degrees
#           
#           # Store the result
#           angles <- rbind(angles, data.table(Atom1 = atom, Atom2 = atom2, Atom3 = atom3, Angle = gamma))
#           
#           print(paste("Calculated Angle:", gamma))
#         }
#       }
#     }
#   }
#   
#   return(angles)
# }
```

### Final Processing

```{r}
# Initialize an empty data table to store the results
main_table <- data.table()

# Iterate over each CIF content
for (cif in cif_tab) {
  # Extract various properties and metrics
  database_code <- extract_database_code(cif)
  chemical_formula <- extract_chemical_formula(cif)
  structure_type <- extract_structure_type(cif)
  space_group_name <- extract_space_group_name(cif)
  space_group_number <- extract_space_group_number(cif)
  unit_cell_metrics <- extract_unit_cell_metrics(cif)
  atomic_coordinates <- extract_atomic_coordinates(cif)
  symmetry_operations <- extract_symmetry_operations(cif)
  transformed_coords <- apply_symmetry_operations(atomic_coordinates, symmetry_operations)
  expanded_coords <- expand_transformed_coords(transformed_coords)
  distances <- calculate_distances(atomic_coordinates, expanded_coords, unit_cell_metrics)

  # Apply the coordination number prediction algorithms
  bonded_pairs_minimum_distance <- minimum_distance(distances)
  bonded_pairs_brunner <- brunner(distances)
  bonded_pairs_hoppe <- hoppe(distances)
  
  # Calculate bond angles for different bonded pairs
  # angles_minimum_distance <- calculate_angles(bonded_pairs_minimum_distance, distances)
  # angles_brunner <- calculate_angles(bonded_pairs_brunner, distances)
  # angles_hoppe <- calculate_angles(bonded_pairs_hoppe, distances)
  

  # Combine extracted data into the main table
  main_table <- rbindlist(list(main_table, data.table(
    database_code = database_code,
    chemical_formula = chemical_formula,
    structure_type = structure_type,
    space_group_name = space_group_name,
    space_group_number = space_group_number,
    unit_cell_metrics = list(unit_cell_metrics),
    atomic_coordinates = list(atomic_coordinates),
    symmetry_operations = list(symmetry_operations),
    transformed_coords = list(transformed_coords),
    expanded_coords = list(expanded_coords),
    distances = list(distances),
    bonded_pairs_minimum_distance = list(bonded_pairs_minimum_distance),
    bonded_pairs_brunner = list(bonded_pairs_brunner),
    bonded_pairs_hoppe = list(bonded_pairs_hoppe)#,
    # angles_minimum_distance = list(angles_minimum_distance),
    # angles_brunner = list(angles_brunner),
    # angles_hoppe = list(angles_hoppe)
  )), fill = TRUE)
}
```
