---
title: "Project Plan"
author: "Don Ngo"
date: "`r Sys.Date()`"
output: pdf_document
---

#Packages (to be adjusted)
```{r}
install.packages("cry")
install.packages("data.table")
library(cry)
library(data.table)
library(readtext)
library(stingr)
```

# General Objectives
- Create a private GitHub repository.
- Extract information from crystallography files.
- Identify necessary information for cluster analyses.
- Build functions.
- Build a package.

# Specific Objectives and Tasks

## Task 1: Create a Private GitHub Repository

**Action:** Set up a private repository on GitHub.  
**Requirement:** Each code segment added should be accompanied by an explanation.

## Task 2: Bind CIF Files into One Table

**Steps:**

1. **Download CIF files from the provided Google Drive link.**

2. **Set the working directory:**

```{r}
setwd('/Users/dngo/repos/ml-crystals')
```

3. **List CIF files**

```{r}
cif_list <- list.files(pattern="\\.cif$", full.names = TRUE)
cif_tab <- lapply(cif_list, readLines)
```

