---
title: "Chemistry-Based Analysis: Sulfides, Selenides, and Tellurides"
author: "crystract Workflow"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
    code_folding: show
    fig_width: 14
    fig_height: 10
  pdf_document:
    toc: true
subtitle: High CN Diagnosis & Weighted CN Analysis
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = TRUE, warning = FALSE)
```

## 1. Introduction

This workflow performs a targeted analysis of Sulfides, Selenides, and Tellurides. It utilizes the Minimum Distance bonding algorithm within the parallelized extraction pipeline and includes diagnostic steps for high Coordination Numbers (CN > 12) and Weighted Coordination Numbers.

**Note:** All plots in this report are interactive. You can hover to see details, zoom in/out, and click legend items to filter data.

## 2. Setup: Libraries and Reference Data

```{r load-libraries, message=FALSE, warning=FALSE}
# Data Manipulation
library(data.table)
library(stringr)
library(tidyr)
library(dplyr)

# Crystallography
library(crystract)

# Parallel Processing
library(future)
library(future.apply)

# Visualization
library(ggplot2)
library(forcats)
library(viridis)
library(gridExtra)
library(DT)      # For interactive tables
library(plotly)  # For interactive plots

# --- Initialize Time Tracking ---
timing_log <- data.frame(
  Step = character(), 
  Duration_Mins = numeric(),
  stringsAsFactors = FALSE
)

log_time <- function(step_name, start_time) {
  duration <- as.numeric(difftime(Sys.time(), start_time, units = "mins"))
  return(data.frame(Step = step_name, Duration_Mins = duration))
}

global_start <- Sys.time()

# --- ATOMIC NUMBER LOOKUP (For Sorting) ---
ptable <- data.table(
  Symbol = c("H","He","Li","Be","B","C","N","O","F","Ne",
             "Na","Mg","Al","Si","P","S","Cl","Ar","K","Ca",
             "Sc","Ti","V","Cr","Mn","Fe","Co","Ni","Cu","Zn",
             "Ga","Ge","As","Se","Br","Kr","Rb","Sr","Y","Zr",
             "Nb","Mo","Tc","Ru","Rh","Pd","Ag","Cd","In","Sn",
             "Sb","Te","I","Xe","Cs","Ba","La","Ce","Pr","Nd",
             "Pm","Sm","Eu","Gd","Tb","Dy","Ho","Er","Tm","Yb",
             "Lu","Hf","Ta","W","Re","Os","Ir","Pt","Au","Hg",
             "Tl","Pb","Bi","Po","At","Rn","Fr","Ra","Ac","Th",
             "Pa","U","Np","Pu","Am","Cm","Bk","Cf","Es","Fm"),
  Z = 1:100
)
```

## 3. Data Discovery and Chemistry Filtering

Filter criteria: **(Contains S OR Se OR Te) AND (Does NOT contain O)**.

```{r filter-by-chemistry}
# --- Configuration ---
# UPDATE THIS PATH to your local CIF directory
cif_directory <- "/Users/carnegieuser/Documents/amcsd" 

step_start <- Sys.time()

# --- 1. Get list of files ---
if (!dir.exists(cif_directory)) {
  warning("The specified directory does not exist. Please update 'cif_directory'.")
  all_cif_files <- character(0)
} else {
  all_cif_files <- list.files(cif_directory, pattern = "\\.cif$", full.names = TRUE)
}

if(length(all_cif_files) == 0) stop("No .cif files found.")

cat(sprintf("Found %d CIF files. Checking chemistry...\n", length(all_cif_files)))

# --- 2. Define Chemistry Filter Logic ---
check_chemistry <- function(file_path) {
  tryCatch({
    cif_content <- read_cif_files(file_path)[[1]]
    formula_str <- extract_chemical_formula(cif_content)
    if (is.na(formula_str)) return(FALSE)
    
    # Extract elements
    elements <- unique(stringr::str_extract_all(formula_str, "[A-Z][a-z]?")[[1]])
    
    # Criteria: Has S/Se/Te AND No Oxygen
    has_target <- any(c("S", "Se", "Te") %in% elements)
    has_oxygen <- "O" %in% elements
    
    return(has_target && !has_oxygen)
  }, error = function(e) return(FALSE))
}

# --- 3. Run Filter (Parallelized) ---
plan(multisession, workers = parallel::detectCores() - 1)
matches_mask <- future_sapply(all_cif_files, check_chemistry)
target_files <- all_cif_files[matches_mask]
plan(sequential)

timing_log <- rbind(timing_log, log_time("1. Filtering", step_start))
cat(sprintf("Filtering complete. Kept: %d files.\n", length(target_files)))
```

## 4. Parallel Analysis (Optimized)

We now run `crystract` with `bonding_algorithms = "minimum_distance"` and `calculate_bond_angles = TRUE` using a customized robust worker function.

```{r run-analysis}
# --- Define Custom Radii Table ---
custom_radii_values <- data.table(
  Symbol = c("Ac", "Ag", "Al", "Am", "Ar", "As", "At", "Au", "B", "Ba", 
             "Be", "Bi", "Br", "C", "Ca", "Cd", "Ce", "Cl", "Cm", "Co", 
             "Cr", "Cs", "Cu", "Dy", "Er", "Eu", "F", "Fe", "Fr", "Ga", 
             "Gd", "Ge", "H", "He", "Hf", "Hg", "Ho", "I", "In", "Ir", 
             "K", "Kr", "La", "Li", "Lu", "Mg", "Mn", "Mo", "N", "Na", 
             "Nb", "Nd", "Ne", "Ni", "Np", "O", "Os", "P", "Pa", "Pb", 
             "Pd", "Pm", "Po", "Pr", "Pt", "Pu", "Ra", "Rb", "Re", "Rh", 
             "Rn", "Ru", "S", "Sb", "Sc", "Se", "Si", "Sm", "Sn", "Sr", 
             "Ta", "Tb", "Tc", "Te", "Th", "Ti", "Tl", "Tm", "U", "V", 
             "W", "Xe", "Y", "Yb", "Zn", "Zr"),
  Radius = c(2.15, 1.45, 1.21, 1.8, 1.06, 1.19, 1.5, 1.36, 0.84, 2.15, 
             0.96, 1.48, 1.2, 0.73, 1.76, 1.44, 2.04, 1.02, 1.69, 1.38, 
             1.39, 2.44, 1.32, 1.92, 1.89, 1.98, 0.57, 1.42, 2.6, 1.22, 
             1.96, 1.2, 0.31, 0.28, 1.75, 1.32, 1.92, 1.39, 1.42, 1.41, 
             2.03, 1.16, 2.07, 1.28, 1.87, 1.41, 1.5, 1.54, 0.71, 1.66, 
             1.64, 2.01, 0.58, 1.24, 1.9, 0.66, 1.44, 1.07, 2, 1.46, 
             1.39, 1.99, 1.4, 2.03, 1.36, 1.87, 2.21, 2.2, 1.51, 1.42, 
             1.5, 1.46, 1.05, 1.39, 1.7, 1.2, 1.11, 1.98, 1.39, 1.95, 
             1.7, 1.94, 1.47, 1.38, 2.06, 1.6, 1.45, 1.9, 1.96, 1.53, 
             1.62, 1.4, 1.9, 1.87, 1.22, 1.75),
  Type = "covalent" 
)

set_radii_data(custom_radii_values)
cat("Custom radii table successfully loaded.\n")

analysis_output_dir <- "~/repos/ml-crystals/notebook/natural/crystract_chemistry_results"
num_workers <- parallel::detectCores() - 1 

# --- Robust Worker Function ---
process_structure_robust <- function(file_path) {
  fname <- basename(file_path)
  
  tryCatch({
    # 1. Read and Extract
    cif_list <- read_cif_files(file_path)
    if(length(cif_list) == 0) return(NULL)
    cif_data <- cif_list[[1]]
    
    formula <- extract_chemical_formula(cif_data)
    atoms   <- extract_atomic_coordinates(cif_data, formula)
    metrics <- extract_unit_cell_metrics(cif_data)
    sym_ops <- extract_symmetry_operations(cif_data)
    
    if (is.null(atoms) || is.null(metrics) || is.null(sym_ops)) return(NULL)
    
    # 2. Transform and Expand (3x3x3 Supercell)
    full_cell <- apply_symmetry_operations(atoms, sym_ops, metrics, tolerance = 1e-4)
    expanded  <- expand_transformed_coords(full_cell, expansion_factors = c(1, 1, 1))
    
    # 3. Calculate Raw Distances
    dists <- calculate_distances(atoms, expanded, metrics, tolerance = 1e-4)
    if (nrow(dists) == 0) return(NULL) 
    
    # --- 4. Attempt Ghost Filtering with Fallback ---
    filter_status <- "applied"
    
    ghost_res <- tryCatch({
      filter_ghost_distances(dists, atoms, margin = 0.15, radii_type = "covalent")
    }, error = function(e) list(kept = data.table()))
    
    clean_dists <- ghost_res$kept
    
    # 5. Determine Bonds
    bonds <- minimum_distance(clean_dists, delta = 0.1)
    if (nrow(bonds) == 0) return(NULL)
    
    # 6. Calculate Properties
    cn_table <- calculate_neighbor_counts(bonds)
    angles   <- calculate_angles(bonds, atoms, expanded, metrics)
    
    # 7. Return Result
    data.table(
      file_name = fname,
      chemical_formula = formula,
      filter_status = filter_status, # Log status for diagnosis
      atomic_coordinates = list(atoms),
      expanded_coords = list(expanded),
      bonds_minimum_distance = list(bonds),
      angles_minimum_distance = list(angles),
      cn_minimum_distance = list(cn_table)
    )
    
  }, error = function(e) {
    return(NULL)
  })
}

# --- Execution Loop ---
step_start <- Sys.time()

if (length(target_files) > 0) {
  if (!dir.exists(analysis_output_dir)) dir.create(analysis_output_dir, recursive = TRUE)
  batch_size  <- 1000
  num_batches <- ceiling(length(target_files) / batch_size)
  
  plan(multisession, workers = num_workers)
  
  for (b in 1:num_batches) {
    idx_start <- (b - 1) * batch_size + 1
    idx_end   <- min(b * batch_size, length(target_files))
    batch_files <- target_files[idx_start:idx_end]
    
    cat(sprintf("Processing Batch %d of %d (%d files)...\n", b, num_batches, length(batch_files)))
    results_list <- future_lapply(batch_files, process_structure_robust, future.seed = TRUE)
    batch_dt <- rbindlist(results_list, fill = TRUE)
    
    if (nrow(batch_dt) > 0) {
      save_path <- file.path(analysis_output_dir, paste0("batch_", b, ".rds"))
      saveRDS(batch_dt, save_path)
      
      # Print filter stats for this batch
      stats <- table(batch_dt$filter_status)
      cat("  Filter Status:", paste(names(stats), stats, sep="=", collapse=", "), "\n")
    }
    rm(results_list, batch_dt); gc()
  }
  plan(sequential)
  timing_log <- rbind(timing_log, log_time("2. Analysis (Robust)", step_start))
} else {
  stop("No files passed the chemistry filter.")
}
```

## 5. Data Processing: Aggregation & Diagnosis

```{r process-and-diagnose}
step_start <- Sys.time()

final_bonds_list <- list()
final_angles_list <- list()
high_cn_debug_log <- list()

if (dir.exists(analysis_output_dir)) {
  cat("Reading batch files and aggregating results...\n")
  batch_files <- list.files(analysis_output_dir, pattern = "^batch_\\d+\\.rds$", full.names = TRUE)
  
  for (bf in batch_files) {
    batch_dt <- readRDS(bf)
    
    for (i in 1:nrow(batch_dt)) {
      current_file <- batch_dt$file_name[i]
      if (is.null(batch_dt$bonds_minimum_distance) || length(batch_dt$bonds_minimum_distance) < i) next
      
      bonded_pairs <- batch_dt$bonds_minimum_distance[[i]]
      angles       <- batch_dt$angles_minimum_distance[[i]]
      coords       <- batch_dt$atomic_coordinates[[i]]
      cn_table     <- batch_dt$cn_minimum_distance[[i]]
      expanded_coords <- batch_dt$expanded_coords[[i]]
      
      if (is.null(bonded_pairs) || nrow(bonded_pairs) == 0 || is.null(coords)) next
      coords <- unique(coords, by = "Label")

      # --- DIAGNOSE HIGH GEOMETRIC CN ---
      high_cn_atoms <- cn_table[CoordinationNumber > 12]
      if (nrow(high_cn_atoms) > 0) {
        bad_bonds <- bonded_pairs[Atom1 %in% high_cn_atoms$Atom]
        exp_coords_ref <- unique(expanded_coords[, .(Label, x2=x_a, y2=y_b, z2=z_c)], by = "Label")
        bad_bonds <- merge(bad_bonds, exp_coords_ref, by.x = "Atom2", by.y = "Label", all.x = TRUE)
        bad_bonds[, `:=`(x1 = x2 + DeltaX, y1 = y2 + DeltaY, z1 = z2 + DeltaZ)]
        bad_bonds[, `:=`(File = current_file, Total_CN = cn_table[match(Atom1, Atom), CoordinationNumber])]
        high_cn_debug_log[[length(high_cn_debug_log) + 1]] <- bad_bonds
      }
      
      # --- COLLECT ANGLES ---
      if (!is.null(angles) && nrow(angles) > 0) {
        angles[, file_name := current_file]
        final_angles_list[[length(final_angles_list) + 1]] <- angles
      }
      
      # --- CALCULATE WEIGHTED CN ---
      bonded_pairs[, NeighborSite := sub("_.*", "", Atom2)]
      site_counts <- bonded_pairs[, .(m_i = .N), by = .(Atom1, NeighborSite)]
      site_counts <- merge(site_counts, coords[, .(Label, Occupancy)], 
                           by.x = "NeighborSite", by.y = "Label", all.x = TRUE)
      site_counts[is.na(Occupancy), Occupancy := 1.0]
      w_cn <- site_counts[, .(CN_Weighted = sum(m_i * Occupancy)), by = .(Atom1)]
      
      bonded_pairs <- merge(bonded_pairs, w_cn, by = "Atom1")
      bonded_pairs[, `:=`(
        file_name = current_file,
        RoundedWeightedCN = round(CN_Weighted),
        CoordinationNumber = cn_table[match(Atom1, Atom), CoordinationNumber],
        Element1 = str_extract(Atom1, "^[A-Za-z]+"),
        Element2 = str_extract(Atom2, "^[A-Za-z]+")
      )]
      final_bonds_list[[length(final_bonds_list) + 1]] <- bonded_pairs
    }
  }
  
  final_bonds_data <- rbindlist(final_bonds_list, fill = TRUE)
  all_angles_data <- rbindlist(final_angles_list, fill = TRUE)
  high_cn_diagnosis <- rbindlist(high_cn_debug_log, fill = TRUE)
  
  final_bonds_data <- merge(final_bonds_data, ptable, by.x="Element1", by.y="Symbol", all.x=TRUE)
  setnames(final_bonds_data, "Z", "Z1")
  final_bonds_data <- merge(final_bonds_data, ptable, by.x="Element2", by.y="Symbol", all.x=TRUE)
  setnames(final_bonds_data, "Z", "Z2")
  
  if(nrow(all_angles_data) > 0) {
    all_angles_data[, Element1 := str_extract(CentralAtom, "^[A-Za-z]+")]
    all_angles_data[, Element2 := str_extract(Neighbor1, "^[A-Za-z]+")]
    all_angles_data <- merge(all_angles_data, ptable, by.x="Element1", by.y="Symbol", all.x=TRUE)
    setnames(all_angles_data, "Z", "Z1")
    all_angles_data <- merge(all_angles_data, ptable, by.x="Element2", by.y="Symbol", all.x=TRUE)
    setnames(all_angles_data, "Z", "Z2")
  }

  timing_log <- rbind(timing_log, log_time("3. Process & Diagnose", step_start))
  
  final_bonds_data <- final_bonds_data[!is.na(Z1) & !is.na(Z2)]
  if(nrow(all_angles_data) > 0) all_angles_data <- all_angles_data[!is.na(Z1) & !is.na(Z2)]
  cat("Processing Complete.\n")
}
```

### 5.1 Diagnostics

```{r show-diagnosis}
if (exists("high_cn_diagnosis") && nrow(high_cn_diagnosis) > 0) {
  datatable(head(high_cn_diagnosis[order(-Total_CN)], 500), 
            options = list(scrollX = TRUE, pageLength = 5))
} else {
  cat("No Geometric CN > 12 found.\n")
}
```

## 5.2 File-Level Weighted CN Summary

This table provides a site-by-site breakdown of both the **Geometric Coordination Number** (calculated by the Minimum Distance algorithm) and the **Weighted Coordination Number** (which accounts for site occupancy/disorder). 

Use the search box to find specific files or the column filters to isolate specific elements or high/low coordination environments.

```{r wcn-file-table}
if (exists("final_bonds_data") && nrow(final_bonds_data) > 0) {
  # Aggregate the bond data to the atom-site level
  wcn_summary <- unique(final_bonds_data[, .(
    File = file_name,
    Site = Atom1,
    Element = Element1,
    `Geometric CN` = CoordinationNumber,
    `Weighted CN` = round(CN_Weighted, 3)
  )])

  # Sort by file name for easier navigation
  wcn_summary <- wcn_summary[order(File, Site)]

  # Display as interactive table
  datatable(
    wcn_summary,
    filter = 'top',
    extensions = 'Buttons',
    options = list(
      pageLength = 10,
      autoWidth = TRUE,
      dom = 'Bfrtip',
      buttons = c('copy', 'csv', 'excel'),
      columnDefs = list(list(className = 'dt-center', targets = "_all"))
    ),
    caption = "Summary of Coordination Numbers per Atomic Site"
  )
} else {
  cat("No bond data available to generate summary table.")
}
```

## 7. Global Visualization (Interactive)

### 7.1 Pairwise Bond Lengths

```{r viz-pairwise-length, fig.width=10, fig.height=10}
pair_len_agg <- final_bonds_data[, .(MeanLength = mean(Distance), Count = .N), 
                                 by = .(Element1, Z1, Element2, Z2)]
pair_len_agg <- pair_len_agg[Count >= 5]
pair_len_agg[, Element1 := fct_reorder(Element1, Z1)]
pair_len_agg[, Element2 := fct_reorder(Element2, Z2)]

p71 <- ggplot(pair_len_agg, aes(x = Element2, y = Element1, 
                                text = paste("Center:", Element1, "<br>Neighbor:", Element2, 
                                             "<br>Avg Dist:", round(MeanLength, 3), "Å",
                                             "<br>Count:", Count))) +
  geom_point(aes(size = Count, color = MeanLength)) +
  scale_color_viridis(option = "magma", direction = -1, name = "Length (Å)") +
  scale_size_continuous(range = c(2, 9), name = "Count") +
  labs(title = "Pairwise Bond Lengths (All Elements)", x = "Neighbor", y = "Central") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

ggplotly(p71, tooltip = "text")
```

### 7.2 Pairwise Bond Angles

```{r viz-pairwise-angle, fig.width=10, fig.height=10}
if (exists("all_angles_data") && nrow(all_angles_data) > 0) {
  df_angles <- copy(all_angles_data)
  if(!"Z1" %in% names(df_angles)) {
     df_angles[, Element1 := str_extract(CentralAtom, "^[A-Za-z]+")]
     df_angles[, Element2 := str_extract(Neighbor1, "^[A-Za-z]+")]
     df_angles <- merge(df_angles, ptable, by.x="Element1", by.y="Symbol", all.x=TRUE)
     setnames(df_angles, "Z", "Z1")
     df_angles <- merge(df_angles, ptable, by.x="Element2", by.y="Symbol", all.x=TRUE)
     setnames(df_angles, "Z", "Z2")
  }
  df_angles <- df_angles[!is.na(Z1) & !is.na(Z2)]
  
  pair_ang_agg <- df_angles[, .(MeanAngle = mean(Angle), Count = .N), 
                            by = .(Element1, Z1, Element2, Z2)]
  pair_ang_agg <- pair_ang_agg[Count >= 5]
  pair_ang_agg[, Element1 := fct_reorder(Element1, Z1)]
  pair_ang_agg[, Element2 := fct_reorder(Element2, Z2)]

  p72 <- ggplot(pair_ang_agg, aes(x = Element2, y = Element1,
                                  text = paste("Center:", Element1, "<br>Neighbor:", Element2,
                                               "<br>Avg Angle:", round(MeanAngle, 1), "°",
                                               "<br>Count:", Count))) +
    geom_point(aes(size = Count, color = MeanAngle)) +
    scale_color_viridis(option = "magma", name = "Angle (°)") +
    scale_size_continuous(range = c(2, 9), name = "Count") +
    labs(title = "Pairwise Bond Angles", x = "Neighbor", y = "Central") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

  ggplotly(p72, tooltip = "text")
} else {
  cat("No bond angle data available.\n")
}
```

## 8. Coordination Number Analysis

### 8.1 Regular CN

```{r viz-regular-cn, fig.width=10, fig.height=10}
reg_cn_agg <- final_bonds_data[, .(MeanLength = mean(Distance), Count = .N), 
                               by = .(Element1, Z1, CoordinationNumber)]
reg_cn_agg <- reg_cn_agg[Count >= 10]
reg_cn_agg[, Element1 := fct_reorder(Element1, Z1)]

p81 <- ggplot(reg_cn_agg, aes(x = factor(CoordinationNumber), y = Element1,
                              text = paste("Element:", Element1, "<br>CN:", CoordinationNumber,
                                           "<br>Avg Dist:", round(MeanLength, 3), "Å",
                                           "<br>Count:", Count))) +
  geom_point(aes(size = Count, color = MeanLength)) +
  scale_color_viridis(option = "magma", direction = -1, name = "Length") +
  scale_size_continuous(range = c(2, 8)) +
  labs(title = "Bond Length vs Regular CN", x = "CN", y = "Element") +
  theme_minimal()

ggplotly(p81, tooltip = "text")
```

### 8.2 Weighted CN

```{r viz-weighted-cn-z, fig.width=10, fig.height=10}
wcn_agg <- final_bonds_data[, .(MeanLength = mean(Distance), Count = .N), 
                            by = .(Element1, Z1, RoundedWeightedCN)]
wcn_agg <- wcn_agg[Count >= 10 & RoundedWeightedCN > 0]
wcn_agg[, Element1 := fct_reorder(Element1, Z1)]

p82 <- ggplot(wcn_agg, aes(x = factor(RoundedWeightedCN), y = Element1,
                           text = paste("Element:", Element1, "<br>W-CN:", RoundedWeightedCN,
                                        "<br>Avg Dist:", round(MeanLength, 3), "Å",
                                        "<br>Count:", Count))) +
  geom_point(aes(size = Count, color = MeanLength)) +
  scale_color_viridis(option = "magma", direction = -1, name = "Length") +
  scale_size_continuous(range = c(2, 8)) +
  labs(title = "Bond Length vs Weighted CN", x = "Weighted CN", y = "Element") +
  theme_minimal()

ggplotly(p82, tooltip = "text")
```

### 8.3 Weighted CN (Sorted by Bond Length)

```{r viz-weighted-cn-len, fig.width=10, fig.height=10}
wcn_agg_len <- copy(wcn_agg)
elem_means <- wcn_agg_len[, .(GlobalMean = mean(MeanLength)), by = Element1]
wcn_agg_len <- merge(wcn_agg_len, elem_means, by="Element1")
wcn_agg_len[, Element1 := fct_reorder(Element1, GlobalMean)]

p83 <- ggplot(wcn_agg_len, aes(x = factor(RoundedWeightedCN), y = Element1,
                               text = paste("Element:", Element1, "<br>W-CN:", RoundedWeightedCN,
                                            "<br>Avg Dist:", round(MeanLength, 3), "Å"))) +
  geom_point(aes(size = Count, color = MeanLength)) +
  scale_color_viridis(option = "magma", direction = -1, name = "Length") +
  scale_size_continuous(range = c(2, 10)) +
  labs(title = "Length vs W-CN (Sorted by Length)", x = "Weighted CN", y = "Element") +
  theme_minimal()

ggplotly(p83, tooltip = "text")
```

## 9. Chalcogen-Specific Analysis

```{r chalcogen-setup}
targets <- c("S", "Se", "Te")
df_center_chalc <- final_bonds_data[Element1 %in% targets]
```

### 9.1 S, Se, Te: Bond Lengths (Interactive Bubble)

**Size:** Number of Observations. **Color:** Mean Distance.

```{r chalc-center-len, fig.width=15, fig.height=5}
agg_center <- df_center_chalc[, .(MeanDist = mean(Distance), N=.N), by=.(Element1, Z1, Element2, Z2)]
agg_center <- agg_center[N >= 3]
agg_center[, Element1 := fct_reorder(Element1, Z1)]
agg_center[, Element2 := fct_reorder(Element2, Z2)]

p91 <- ggplot(agg_center, aes(x = Element2, y = Element1,
                              text = paste("Center:", Element1, "<br>Neighbor:", Element2,
                                           "<br>Dist:", round(MeanDist, 3), "Å",
                                           "<br>Obs:", N))) +
  geom_point(aes(size = N, color = MeanDist)) +
  scale_color_viridis(option = "magma", direction = -1, name = "Dist (Å)") +
  scale_size_continuous(range = c(3, 10), name = "Count") +
  labs(title = "Bond Lengths: S, Se, Te Centers", x = "Neighbor", y = "Central") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

ggplotly(p91, tooltip = "text")

# Heatmap
ggplot(agg_center, aes(x = Element2, y = Element1, fill = MeanDist)) +
  geom_tile(color = "white") +
  scale_fill_viridis(option = "magma", direction = -1, name = "Dist (Å)") +
  labs(title = "Bond Lengths: S, Se, Te Centers", 
       subtitle = "Sorted by Atomic Number",
       x = "Neighbor Element", y = "Central Chalcogen") +
  theme_minimal() +
  coord_fixed()
```

### 9.2 S, Se, Te: Regular CN

**Size:** Number of Observations. **Color:** Mean Geometric CN.

```{r chalc-center-cn, fig.width=15, fig.height=5}
agg_center_cn <- df_center_chalc[, .(MeanCN = mean(CoordinationNumber), N=.N), 
                                 by=.(Element1, Z1, Element2, Z2)]
agg_center_cn <- agg_center_cn[N >= 3]
agg_center_cn[, Element1 := fct_reorder(Element1, Z1)]
agg_center_cn[, Element2 := fct_reorder(Element2, Z2)]

p92 <- ggplot(agg_center_cn, aes(x = Element2, y = Element1,
                                 text = paste("Center:", Element1, "<br>Neighbor:", Element2,
                                              "<br>Avg CN:", round(MeanCN, 1),
                                              "<br>Obs:", N))) +
  geom_point(aes(size = N, color = MeanCN)) +
  scale_color_viridis(option = "magma", direction = -1, name = "Avg CN") +
  scale_size_continuous(range = c(3, 10), name = "Count") +
  labs(title = "Regular CN: S, Se, Te Centers", x = "Neighbor", y = "Central") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

ggplotly(p92, tooltip = "text")

# Heatmap
ggplot(agg_center_cn, aes(x = Element2, y = Element1, fill = MeanCN)) +
  geom_tile(color = "white") +
  scale_fill_viridis(option = "magma", direction = -1, name = "Avg CN") +
  labs(title = "Regular Coordination Number: S, Se, Te Centers", 
       subtitle = "Sorted by Atomic Number",
       x = "Neighbor Element", y = "Central Chalcogen") +
  theme_minimal() +
  coord_fixed()
```

### 9.3 S, Se, Te: Weighted CN (Special Aggregation)

**Size:** Average Total CN (Environment size).  
**Color:** Average Specific Weighted CN (Contribution of Neighbor).

```{r chalc-center-wcn, fig.width=15, fig.height=5}
# 1. Calculate Total Weighted CN per Central Atom
total_wcn_per_atom <- final_bonds_data[, .(Total_WCN = sum(CN_Weighted)), by = .(file_name, Atom1)]
plot_data_wcn <- merge(final_bonds_data, total_wcn_per_atom, by = c("file_name", "Atom1"))

# 2. Filter & Aggregate
df_plot <- plot_data_wcn[Element1 %in% targets]
agg_wcn_complex <- df_plot[, .(
  Avg_Specific_WCN = mean(CN_Weighted), # Color
  Avg_Total_WCN    = mean(Total_WCN),   # Size
  N                = .N
), by = .(Element1, Z1, Element2, Z2)]

agg_wcn_complex <- agg_wcn_complex[N >= 3]
agg_wcn_complex[, Element1 := fct_reorder(Element1, Z1)]
agg_wcn_complex[, Element2 := fct_reorder(Element2, Z2)]

p93 <- ggplot(agg_wcn_complex, aes(x = Element2, y = Element1,
                                   text = paste("Center:", Element1, "<br>Neighbor:", Element2,
                                                "<br>Specific WCN (Color):", round(Avg_Specific_WCN, 2),
                                                "<br>Total Env CN (Size):", round(Avg_Total_WCN, 2),
                                                "<br>Obs:", N))) +
  geom_point(aes(size = Avg_Total_WCN, color = Avg_Specific_WCN)) +
  scale_color_viridis(option = "magma", direction = -1, name = "Specific WCN") +
  scale_size_continuous(range = c(2, 9), name = "Total WCN") +
  labs(title = "Weighted CN Analysis", x = "Neighbor", y = "Central") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

ggplotly(p93, tooltip = "text")

# Aggregate Mean Weighted Coordination Number
agg_center_wcn <- df_center_chalc[, .(MeanWCN = mean(CN_Weighted), N=.N), 
                                  by=.(Element1, Z1, Element2, Z2)]
agg_center_wcn <- agg_center_wcn[N >= 3]

# Sort by Atomic Number (Z)
agg_center_wcn[, Element1 := fct_reorder(Element1, Z1)]
agg_center_wcn[, Element2 := fct_reorder(Element2, Z2)]

ggplot(agg_center_wcn, aes(x = Element2, y = Element1, fill = MeanWCN)) +
  geom_tile(color = "white") +
  scale_fill_viridis(option = "magma", direction = -1, name = "W-CN") +
  labs(title = "Weighted Coordination Number: S, Se, Te Centers", 
       subtitle = "Sorted by Atomic Number (Disorder Corrected)",
       x = "Neighbor Element", y = "Central Chalcogen") +
  theme_minimal() +
  coord_fixed()
```

## 10. Distribution Analysis

This visualization compares the distribution of bond lengths for the most common coordination numbers (e.g., 4 and 6), allowing us to see the “spread” of bond distances for specific elements.

```{r viz-distributions, fig.width=12, fig.height=6}
# --- Distribution Analysis (Violin Plot) ---

# 1. Filter for common Coordination Numbers to make the plot readable
common_cns <- c(4, 6) # Focusing on Tetrahedral and Octahedral
violin_data <- final_bonds_data[CoordinationNumber %in% common_cns]

# 2. Filter to top 20 most frequent elements to keep the x-axis clean
top_elements <- violin_data[, .N, by = Element1][order(-N)][1:20, Element1]
violin_data <- violin_data[Element1 %in% top_elements]

# 3. Plot
ggplot(violin_data, aes(x = reorder(Element1, Distance), y = Distance)) +
  geom_violin(aes(fill = factor(CoordinationNumber)), scale = "width", alpha = 0.7, position = position_dodge(width = 0.8)) +
  geom_boxplot(aes(group = interaction(Element1, CoordinationNumber)), 
               width = 0.1, outlier.shape = NA, fill="white", position = position_dodge(width = 0.8)) +
  scale_fill_brewer(palette = "Dark2", name = "CN") +
  labs(title = "Distribution of Bond Lengths",
       subtitle = "Comparing Coordination Numbers 4 and 6 for Top Elements",
       x = "Element (Sorted by Median Distance)",
       y = "Bond Distance (Å)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## 11. Advanced: Length vs Angle (Grouped by Neighbor Family)

Faceted by S, Se, Te. Neighbors colored by Periodic Group (Expanded).

```{r viz-advanced-scatter, fig.width=12, fig.height=6, warning=FALSE, message=FALSE}
# --- 1. Expanded Group Metadata with Split Transition Metals ---
extended_groups <- data.table(
  Symbol = c(
    # Alkali / Alkaline (Groups 1-2)
    "Li","Na","K","Rb","Cs","Fr","Be","Mg","Ca","Sr","Ba","Ra",
    
    # Early Transition Metals (Groups 3-6) - Sc, Ti, V, Cr columns
    "Sc","Y","La","Ti","Zr","Hf","V","Nb","Ta","Cr","Mo","W",
    
    # Late Transition Metals (Groups 7-10) - Mn, Fe, Co, Ni columns
    "Mn","Tc","Re","Fe","Ru","Os","Co","Rh","Ir","Ni","Pd","Pt",
    
    # Coinage & Zinc Groups (Groups 11-12) - Cu, Zn columns
    "Cu","Ag","Au","Zn","Cd","Hg",
    
    # Groups 13-17
    "B","Al","Ga","In","Tl","C","Si","Ge","Sn","Pb","N","P","As",
    "Sb","Bi","O","S","Se","Te","Po","F","Cl","Br","I"
  ),
  GroupLabel = c(
    rep("Alkali/Alkaline (Gr 1-2)", 12),
    rep("Early TMs (Gr 3-6)", 12),
    rep("Late TMs (Gr 7-10)", 12),
    rep("Coinage/Zinc (Gr 11-12)", 6),
    rep("Group 13", 5),
    rep("Group 14", 5),
    rep("Group 15", 5),
    rep("Group 16", 5),
    rep("Group 17", 4)
  )
)

if (exists("all_angles_data") && nrow(all_angles_data) > 0) {
  # A. Filter Angles for S, Se, Te centers ONLY
  target_centers <- c("S", "Se", "Te")
  # Filter using regex to catch isotopes/labels but stricter check later
  df_angles_sub <- all_angles_data[CentralAtom %like% "^(S|Se|Te)([0-9]|$|_)"]
  
  # Extract Base Symbols
  df_angles_sub[, CenterSym := str_extract(CentralAtom, "^[A-Za-z]+")]
  df_angles_sub[, NeighborSym := str_extract(Neighbor1, "^[A-Za-z]+")]
  
  # Strict filter to keep only S, Se, Te (removes false positives like Sc, Sn)
  df_angles_sub <- df_angles_sub[CenterSym %in% target_centers]
  
  # B. Prepare Bonds for Distance merging
  unique_bonds <- unique(final_bonds_data[, .(file_name, Atom1, Atom2, Distance)], by = c("file_name", "Atom1", "Atom2"))
  
  # C. Merge Angle + Distance
  plot_data <- merge(
    df_angles_sub,
    unique_bonds,
    by.x = c("file_name", "CentralAtom", "Neighbor1"),
    by.y = c("file_name", "Atom1", "Atom2"),
    allow.cartesian = TRUE
  )
  
  # D. Merge Neighbor Group Info
  plot_data <- merge(
    plot_data,
    extended_groups,
    by.x = "NeighborSym",
    by.y = "Symbol",
    all.x = TRUE
  )
  
  # Remove NAs (elements not in our group list, e.g., Lanthanides if not mapped)
  plot_data <- plot_data[!is.na(GroupLabel)]
  
  # E. Sort Central Atom for Plotting Order
  plot_data[, CenterSym := factor(CenterSym, levels = c("S", "Se", "Te"))]
  
  # --- 3. Plotting ---
  p11 <- ggplot(plot_data, aes(
    x = Distance,
    y = Angle,
    text = paste(
      "Center:", CentralAtom,
      "<br>Neighbor:", Neighbor1,
      "<br>Group:", GroupLabel
    )
  )) +
    geom_point(aes(color = GroupLabel), alpha = 0.4, size = 1) +
    facet_wrap( ~ CenterSym, scales = "free", ncol = 3) +
    scale_color_viridis_d(option = "turbo", name = "Neighbor Family") +
    labs(
      title = "Bond Length vs Angle Correlation",
      subtitle = "Faceted by Central Atom (S, Se, Te)",
      x = "Bond Length (Å)",
      y = "Bond Angle (°)"
    ) +
    theme_minimal() +
    theme(legend.position = "bottom") +
    guides(color = guide_legend(override.aes = list(alpha = 1, size = 3)))
  
  # Use toWebGL for performance with many points
  toWebGL(ggplotly(p11, tooltip = "text"))
  
} else {
  cat("No angle data available to plot.\n")
}
```