---
title: "Chemistry-Based Analysis: Sulfides, Selenides, and Tellurides"
author: "crystract Workflow"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
    code_folding: show
    fig_width: 14
    fig_height: 10
  pdf_document:
    toc: true
subtitle: ECon NN Diagnosis & Weighted CN Analysis
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = TRUE, warning = FALSE)
```

## 1. Introduction

This workflow performs a targeted, large-scale crystallographic analysis of Sulfides, Selenides, and Tellurides. By utilizing the **ECon (Effective Coordination Number)** bonding algorithm within the highly parallelized `crystract` extraction pipeline, we calculate fundamental geometric metrics such as bond distances, normalized bond distances (divided by the sum of atomic radii), bond angles, and coordination numbers (CN) across thousands of parsed `.cif` files.

This notebook includes both the baseline **ECon Coordination Number** (which evaluates the geometric and effective environment without occupancy adjustments) and the **Weighted Coordination Number (WCN)**. Weighted CN provides a secondary, critical representation of the local chemical environment by taking into account crystallographic site occupancies to resolve the partial occupancies inherent to structural disorder. Both are provided for a comprehensive comparison, with a special focus on S, Se, and Te.

## 2. Setup: Libraries and Reference Data

We begin by loading the necessary data manipulation, crystallography, and parallel processing packages. A robust periodic table reference (`ptable`) and custom covalent radii are established to ensure accurate normalization and visualization.

```{r load-libraries, message=FALSE, warning=FALSE}
# Data Manipulation
library(data.table)
library(stringr)
library(tidyr)
library(dplyr)

# Crystallography
library(crystract)

# Parallel Processing
library(future)
library(future.apply)

# Visualization
library(ggplot2)
library(forcats)
library(viridis)
library(gridExtra)
library(DT)      # For interactive tables
library(plotly)  # For interactive plots
library(scales)  # For plot scales and breaks

# --- Initialize Time Tracking ---
timing_log <- data.frame(
  Step = character(), 
  Duration_Mins = numeric(),
  stringsAsFactors = FALSE
)

log_time <- function(step_name, start_time) {
  duration <- as.numeric(difftime(Sys.time(), start_time, units = "mins"))
  return(data.frame(Step = step_name, Duration_Mins = duration))
}

global_start <- Sys.time()

# --- ATOMIC NUMBER LOOKUP (For Sorting) ---
ptable <- data.table(
  Symbol = c("H","He","Li","Be","B","C","N","O","F","Ne",
             "Na","Mg","Al","Si","P","S","Cl","Ar","K","Ca",
             "Sc","Ti","V","Cr","Mn","Fe","Co","Ni","Cu","Zn",
             "Ga","Ge","As","Se","Br","Kr","Rb","Sr","Y","Zr",
             "Nb","Mo","Tc","Ru","Rh","Pd","Ag","Cd","In","Sn",
             "Sb","Te","I","Xe","Cs","Ba","La","Ce","Pr","Nd",
             "Pm","Sm","Eu","Gd","Tb","Dy","Ho","Er","Tm","Yb",
             "Lu","Hf","Ta","W","Re","Os","Ir","Pt","Au","Hg",
             "Tl","Pb","Bi","Po","At","Rn","Fr","Ra","Ac","Th",
             "Pa","U","Np","Pu","Am","Cm","Bk","Cf","Es","Fm"),
  Z = 1:100
)

# --- Define Custom Radii Table ---
custom_radii_values <- data.table(
  Symbol = c("Ac", "Ag", "Al", "Am", "Ar", "As", "At", "Au", "B", "Ba", 
             "Be", "Bi", "Br", "C", "Ca", "Cd", "Ce", "Cl", "Cm", "Co", 
             "Cr", "Cs", "Cu", "Dy", "Er", "Eu", "F", "Fe", "Fr", "Ga", 
             "Gd", "Ge", "H", "He", "Hf", "Hg", "Ho", "I", "In", "Ir", 
             "K", "Kr", "La", "Li", "Lu", "Mg", "Mn", "Mo", "N", "Na", 
             "Nb", "Nd", "Ne", "Ni", "Np", "O", "Os", "P", "Pa", "Pb", 
             "Pd", "Pm", "Po", "Pr", "Pt", "Pu", "Ra", "Rb", "Re", "Rh", 
             "Rn", "Ru", "S", "Sb", "Sc", "Se", "Si", "Sm", "Sn", "Sr", 
             "Ta", "Tb", "Tc", "Te", "Th", "Ti", "Tl", "Tm", "U", "V", 
             "W", "Xe", "Y", "Yb", "Zn", "Zr"),
  Radius = c(2.15, 1.45, 1.21, 1.8, 1.06, 1.19, 1.5, 1.36, 0.84, 2.15, 
             0.96, 1.48, 1.2, 0.73, 1.76, 1.44, 2.04, 1.02, 1.69, 1.38, 
             1.39, 2.44, 1.32, 1.92, 1.89, 1.98, 0.57, 1.42, 2.6, 1.22, 
             1.96, 1.2, 0.31, 0.28, 1.75, 1.32, 1.92, 1.39, 1.42, 1.41, 
             2.03, 1.16, 2.07, 1.28, 1.87, 1.41, 1.5, 1.54, 0.71, 1.66, 
             1.64, 2.01, 0.58, 1.24, 1.9, 0.66, 1.44, 1.07, 2, 1.46, 
             1.39, 1.99, 1.4, 2.03, 1.36, 1.87, 2.21, 2.2, 1.51, 1.42, 
             1.5, 1.46, 1.05, 1.39, 1.7, 1.2, 1.11, 1.98, 1.39, 1.95, 
             1.7, 1.94, 1.47, 1.38, 2.06, 1.6, 1.45, 1.9, 1.96, 1.53, 
             1.62, 1.4, 1.9, 1.87, 1.22, 1.75),
  Type = "covalent" 
)
```

## 3. Data Discovery and Chemistry Filtering

To analyze Chalcogenides strictly, our filtering logic is: **(Contains S OR Se OR Te) AND (Does NOT contain O)**. Excluding oxygen prevents the inclusion of sulfates, selenites, and tellurates, isolating the data strictly to sulfides, selenides, and tellurides.

```{r filter-by-chemistry}
# --- Configuration ---
cif_directory <- "/Users/donngo/Documents/cif" 

step_start <- Sys.time()

# --- 1. Get list of files ---
if (!dir.exists(cif_directory)) {
  warning("The specified directory does not exist. Please update 'cif_directory'.")
  all_cif_files <- character(0)
} else {
  all_cif_files <- list.files(cif_directory, pattern = "\\.cif$", full.names = TRUE)
}

if(length(all_cif_files) == 0) stop("No .cif files found.")

cat(sprintf("Found %d CIF files. Checking chemistry...\n", length(all_cif_files)))

# --- 2. Define Chemistry Filter Logic ---
check_chemistry <- function(file_path) {
  tryCatch({
    cif_content <- read_cif_files(file_path)[[1]]
    formula_str <- extract_chemical_formula(cif_content)
    if (is.na(formula_str)) return(FALSE)
    
    # Extract elements
    elements <- unique(stringr::str_extract_all(formula_str, "[A-Z][a-z]?")[[1]])
    
    # Criteria: Has S/Se/Te AND No Oxygen
    has_target <- any(c("S", "Se", "Te") %in% elements)
    has_oxygen <- "O" %in% elements
    
    return(has_target && !has_oxygen)
  }, error = function(e) return(FALSE))
}

# --- 3. Run Filter (Parallelized) ---
plan(multisession, workers = 5)
matches_mask <- future_sapply(all_cif_files, check_chemistry)
target_files <- all_cif_files[matches_mask]
plan(sequential)

timing_log <- rbind(timing_log, log_time("1. Filtering", step_start))
cat(sprintf("Filtering complete. Kept: %d files.\n", length(target_files)))
```

## 4. Parallel Analysis (ECon Algorithm)

We run the `crystract` feature extraction using the `econ` bonding algorithm. This chunk expands the unit cell into a 3x3x3 supercell to ensure edge-atoms are correctly bonded to their periodic neighbors. A ghost filtering mechanism is also applied to handle duplicate distance measurements often created by crystallographic disorder.

```{r run-analysis}
set_radii_data(custom_radii_values)
cat("Custom radii table successfully loaded for filtering.\n")

analysis_output_dir <- "~/repos/ml-crystals/notebook/natural/crystract_econ_results"
num_workers <- parallel::detectCores() - 1 

# --- Robust Worker Function ---
process_structure_robust <- function(file_path) {
  fname <- basename(file_path)
  
  tryCatch({
    # 1. Read and Extract
    cif_list <- read_cif_files(file_path)
    if(length(cif_list) == 0) return(NULL)
    cif_data <- cif_list[[1]]
    
    formula <- extract_chemical_formula(cif_data)
    atoms   <- extract_atomic_coordinates(cif_data, formula)
    metrics <- extract_unit_cell_metrics(cif_data)
    sym_ops <- extract_symmetry_operations(cif_data)
    
    if (is.null(atoms) || is.null(metrics) || is.null(sym_ops)) return(NULL)
    
    # 2. Transform and Expand (3x3x3 Supercell)
    full_cell <- apply_symmetry_operations(atoms, sym_ops, metrics, tolerance = 1e-4)
    expanded  <- expand_transformed_coords(full_cell, expansion_factors = c(1, 1, 1))
    
    # 3. Calculate Raw Distances
    dists <- calculate_distances(atoms, expanded, metrics, tolerance = 1e-4)
    if (nrow(dists) == 0) return(NULL) 
    
    # --- 4. Attempt Ghost Filtering ---
    filter_status <- "applied"
    
    ghost_res <- tryCatch({
      filter_ghost_distances(dists, atoms, margin = 0.15, radii_type = "covalent")
    }, error = function(e) list(kept = data.table()))
    
    clean_dists <- ghost_res$kept
    
    # 5. Determine Bonds using ECon
    bonds <- econ(clean_dists)
    if (nrow(bonds) == 0) return(NULL)
    
    # 6. Calculate Properties
    cn_table <- calculate_neighbor_counts(bonds)
    angles   <- calculate_angles(bonds, atoms, expanded, metrics)
    
    # 7. Return Result
    data.table(
      file_name = fname,
      chemical_formula = formula,
      filter_status = filter_status, # Log status for diagnosis
      atomic_coordinates = list(atoms),
      expanded_coords = list(expanded),
      bonds_econ = list(bonds),
      angles_econ = list(angles),
      cn_econ = list(cn_table)
    )
    
  }, error = function(e) {
    return(NULL)
  })
}

# --- Execution Loop ---
step_start <- Sys.time()

if (length(target_files) > 0) {
  if (!dir.exists(analysis_output_dir)) dir.create(analysis_output_dir, recursive = TRUE)
  batch_size  <- 500
  num_batches <- ceiling(length(target_files) / batch_size)
  
  plan(multisession, workers = num_workers)
  
  for (b in 1:num_batches) {
    idx_start <- (b - 1) * batch_size + 1
    idx_end   <- min(b * batch_size, length(target_files))
    batch_files <- target_files[idx_start:idx_end]
    
    cat(sprintf("Processing Batch %d of %d (%d files)...\n", b, num_batches, length(batch_files)))
    results_list <- future_lapply(batch_files, process_structure_robust, future.seed = TRUE)
    batch_dt <- rbindlist(results_list, fill = TRUE)
    
    if (nrow(batch_dt) > 0) {
      save_path <- file.path(analysis_output_dir, paste0("batch_", b, ".rds"))
      saveRDS(batch_dt, save_path)
      
      # Print filter stats for this batch
      stats <- table(batch_dt$filter_status)
      cat("  Filter Status:", paste(names(stats), stats, sep="=", collapse=", "), "\n")
    }
    rm(results_list, batch_dt); gc()
  }
  plan(sequential)
  timing_log <- rbind(timing_log, log_time("2. Analysis (Robust)", step_start))
} else {
  stop("No files passed the chemistry filter.")
}
```

## 5. Data Processing: Aggregation & Diagnosis

In this step, we collect the ECon output and map elemental data (Atomic Number, Covalent Radii) to the central and neighbor atoms. We compute the **Normalized Distance** for each bond by dividing the geometric distance by the sum of the covalent radii of the two bonded elements. Additionally, we compute the **Weighted Coordination Number (WCN)** by weighing the geometric instances against structural occupancy.

```{r process-and-diagnose}
step_start <- Sys.time()

final_bonds_list <- list()
final_angles_list <- list()
high_cn_debug_log <- list()

if (dir.exists(analysis_output_dir)) {
  cat("Reading batch files and aggregating results...\n")
  batch_files <- list.files(analysis_output_dir, pattern = "^batch_\\d+\\.rds$", full.names = TRUE)
  
  for (bf in batch_files) {
    batch_dt <- readRDS(bf)
    
    for (i in 1:nrow(batch_dt)) {
      current_file <- batch_dt$file_name[i]
      if (is.null(batch_dt$bonds_econ) || length(batch_dt$bonds_econ) < i) next
      
      bonded_pairs <- batch_dt$bonds_econ[[i]]
      angles       <- batch_dt$angles_econ[[i]]
      coords       <- batch_dt$atomic_coordinates[[i]]
      cn_table     <- batch_dt$cn_econ[[i]]
      expanded_coords <- batch_dt$expanded_coords[[i]]
      
      if (is.null(bonded_pairs) || nrow(bonded_pairs) == 0 || is.null(coords)) next
      coords <- unique(coords, by = "Label")

      # --- DIAGNOSE HIGH GEOMETRIC CN ---
      high_cn_atoms <- cn_table[CoordinationNumber > 12]
      if (nrow(high_cn_atoms) > 0) {
        bad_bonds <- bonded_pairs[Atom1 %in% high_cn_atoms$Atom]
        exp_coords_ref <- unique(expanded_coords[, .(Label, x2=x_a, y2=y_b, z2=z_c)], by = "Label")
        bad_bonds <- merge(bad_bonds, exp_coords_ref, by.x = "Atom2", by.y = "Label", all.x = TRUE)
        bad_bonds[, `:=`(x1 = x2 + DeltaX, y1 = y2 + DeltaY, z1 = z2 + DeltaZ)]
        bad_bonds[, `:=`(File = current_file, Total_CN = cn_table[match(Atom1, Atom), CoordinationNumber])]
        high_cn_debug_log[[length(high_cn_debug_log) + 1]] <- bad_bonds
      }
      
      # --- COLLECT ANGLES ---
      if (!is.null(angles) && nrow(angles) > 0) {
        angles[, file_name := current_file]
        final_angles_list[[length(final_angles_list) + 1]] <- angles
      }
      
      # --- CALCULATE WEIGHTED CN ---
      bonded_pairs[, NeighborSite := sub("_.*", "", Atom2)]
      site_counts <- bonded_pairs[, .(m_i = .N), by = .(Atom1, NeighborSite)]
      site_counts <- merge(site_counts, coords[, .(Label, Occupancy)], 
                           by.x = "NeighborSite", by.y = "Label", all.x = TRUE)
      site_counts[is.na(Occupancy), Occupancy := 1.0]
      w_cn <- site_counts[, .(CN_Weighted = sum(m_i * Occupancy)), by = .(Atom1)]
      
      # --- ASSEMBLE DATA ---
      # Merge Exact Neighbor Occupancies into final dataset for Specific WCN plots
      bonded_pairs <- merge(bonded_pairs, coords[, .(Label, Occupancy)], by.x = "NeighborSite", by.y = "Label", all.x = TRUE)
      bonded_pairs[is.na(Occupancy), Occupancy := 1.0]
      
      bonded_pairs <- merge(bonded_pairs, w_cn, by = "Atom1")
      bonded_pairs[, `:=`(
        file_name = current_file,
        CoordinationNumber = cn_table[match(Atom1, Atom), CoordinationNumber],
        RoundedWeightedCN = round(CN_Weighted),
        Element1 = str_extract(Atom1, "^[A-Za-z]+"),
        Element2 = str_extract(Atom2, "^[A-Za-z]+")
      )]
      final_bonds_list[[length(final_bonds_list) + 1]] <- bonded_pairs
    }
  }
  
  final_bonds_data <- rbindlist(final_bonds_list, fill = TRUE)
  all_angles_data <- rbindlist(final_angles_list, fill = TRUE)
  high_cn_diagnosis <- rbindlist(high_cn_debug_log, fill = TRUE)
  
  # Merge Atomic Numbers
  final_bonds_data <- merge(final_bonds_data, ptable, by.x="Element1", by.y="Symbol", all.x=TRUE)
  setnames(final_bonds_data, "Z", "Z1")
  final_bonds_data <- merge(final_bonds_data, ptable, by.x="Element2", by.y="Symbol", all.x=TRUE)
  setnames(final_bonds_data, "Z", "Z2")
  
  # Merge Custom Radii for Normalized Distance Calculation
  final_bonds_data <- merge(final_bonds_data, custom_radii_values[, .(Symbol, Radius)], by.x="Element1", by.y="Symbol", all.x=TRUE)
  setnames(final_bonds_data, "Radius", "Radius1")
  final_bonds_data <- merge(final_bonds_data, custom_radii_values[, .(Symbol, Radius)], by.x="Element2", by.y="Symbol", all.x=TRUE)
  setnames(final_bonds_data, "Radius", "Radius2")
  
  # Calculate Normalized Distance = Distance / (Radius1 + Radius2)
  final_bonds_data[, NormalizedDistance := Distance / (Radius1 + Radius2)]
  
  if(nrow(all_angles_data) > 0) {
    all_angles_data[, Element1 := str_extract(CentralAtom, "^[A-Za-z]+")]
    all_angles_data[, Element2 := str_extract(Neighbor1, "^[A-Za-z]+")]
    all_angles_data <- merge(all_angles_data, ptable, by.x="Element1", by.y="Symbol", all.x=TRUE)
    setnames(all_angles_data, "Z", "Z1")
    all_angles_data <- merge(all_angles_data, ptable, by.x="Element2", by.y="Symbol", all.x=TRUE)
    setnames(all_angles_data, "Z", "Z2")
  }

  timing_log <- rbind(timing_log, log_time("3. Process & Diagnose", step_start))
  
  final_bonds_data <- final_bonds_data[!is.na(Z1) & !is.na(Z2)]
  if(nrow(all_angles_data) > 0) all_angles_data <- all_angles_data[!is.na(Z1) & !is.na(Z2)]
  cat("Processing Complete.\n")
}
```

### 5.1 Diagnostics

This table allows us to peek into files where elements exhibited mathematically massive coordination numbers. Extremely high ECon CNs (> 12) usually indicate deeply disordered structures, overlapping atomic positions, or parsing errors.

```{r show-diagnosis}
if (exists("high_cn_diagnosis") && nrow(high_cn_diagnosis) > 0) {
  datatable(head(high_cn_diagnosis[order(-Total_CN)], 500), 
            options = list(scrollX = TRUE, pageLength = 5))
} else {
  cat("No ECon CN > 12 found.\n")
}
```

## 5.2 File-Level Coordination Summary

This table provides a site-by-site breakdown of the **ECon Coordination Number** alongside the occupancy-corrected **Weighted CN**. Use the search box to find specific files or the column filters to isolate specific elements.

```{r cn-file-table}
if (exists("final_bonds_data") && nrow(final_bonds_data) > 0) {
  # Aggregate the bond data to the atom-site level
  cn_summary <- unique(final_bonds_data[, .(
    File = file_name,
    Site = Atom1,
    Element = Element1,
    `ECon CN` = round(CoordinationNumber, 3),
    `Weighted CN` = round(CN_Weighted, 3)
  )])

  # Sort by file name for easier navigation
  cn_summary <- cn_summary[order(File, Site)]

  # Display as interactive table
  datatable(
    cn_summary,
    filter = 'top',
    extensions = 'Buttons',
    options = list(
      pageLength = 10,
      autoWidth = TRUE,
      dom = 'Bfrtip',
      buttons = c('copy', 'csv', 'excel'),
      columnDefs = list(list(className = 'dt-center', targets = "_all"))
    ),
    caption = "Summary of ECon and Weighted Coordination Numbers per Atomic Site"
  )
} else {
  cat("No bond data available to generate summary table.")
}
```

---

**A Note on the Visualizations:**
Because interactive web-based plotting engines (like `plotly`) frequently suppress or hide size mapping legends, a dual-output strategy is implemented below for all bubble plots. 
1. **Static Export Image:** The first output is a standard `ggplot2` image. It correctly renders both the color scale and the bubble size scale, making it the perfect target to save or export for reports.
2. **Interactive Analysis Tool:** The second output is an interactive `plotly` frame. While the size legend might be hidden, this plot enables you to hover over specific bubbles to query their exact underlying metrics (count, distance, etc.) and zoom into dense regions.

## 7. Global Visualization

### 7.1 Pairwise Bond Lengths

This plot correlates the central atom against all possible neighbor atoms found across the database. 
* **Bubble Color:** Indicates the mean bond length (Å) for that element pairing.
* **Bubble Size:** Indicates the total number of distinct observations for that pairing.

```{r viz-pairwise-length, fig.width=10, fig.height=10}
pair_len_agg <- final_bonds_data[, .(MeanLength = mean(Distance), Count = .N), 
                                 by = .(Element1, Z1, Element2, Z2)]
pair_len_agg <- pair_len_agg[Count >= 5]
pair_len_agg[, Element1 := fct_reorder(Element1, Z1)]
pair_len_agg[, Element2 := fct_reorder(Element2, Z2)]

p71 <- ggplot(pair_len_agg, aes(x = Element2, y = Element1, 
                                text = paste("Center:", Element1, "<br>Neighbor:", Element2, 
                                             "<br>Avg Dist:", round(MeanLength, 3), "Å",
                                             "<br>Count:", Count))) +
  geom_point(aes(size = Count, color = MeanLength)) +
  scale_color_viridis_c(option = "magma", direction = -1, name = "Length (Å)") +
  scale_size_continuous(range = c(2, 9), breaks = scales::breaks_pretty(n = 5), name = "Count (Observations)") +
  labs(title = "Pairwise Bond Lengths (All Elements)", x = "Neighbor Element", y = "Central Element") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

# 1. Static output for export (Preserves legend)
print(p71)

# 2. Interactive output for exploration
ggplotly(p71, tooltip = "text")
```

### 7.2 Pairwise Bond Angles

```{r viz-pairwise-angle, fig.width=10, fig.height=10}
if (exists("all_angles_data") && nrow(all_angles_data) > 0) {
  df_angles <- copy(all_angles_data)
  
  pair_ang_agg <- df_angles[, .(MeanAngle = mean(Angle), Count = .N), 
                            by = .(Element1, Z1, Element2, Z2)]
  pair_ang_agg <- pair_ang_agg[Count >= 5]
  pair_ang_agg[, Element1 := fct_reorder(Element1, Z1)]
  pair_ang_agg[, Element2 := fct_reorder(Element2, Z2)]

  p72 <- ggplot(pair_ang_agg, aes(x = Element2, y = Element1,
                                  text = paste("Center:", Element1, "<br>Neighbor:", Element2,
                                               "<br>Avg Angle:", round(MeanAngle, 1), "°",
                                               "<br>Count:", Count))) +
    geom_point(aes(size = Count, color = MeanAngle)) +
    scale_color_viridis_c(option = "magma", name = "Angle (°)") +
    scale_size_continuous(range = c(2, 9), breaks = scales::breaks_pretty(n = 5), name = "Count (Observations)") +
    labs(title = "Pairwise Bond Angles", x = "Neighbor Element", y = "Central Element") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

  # 1. Static output for export (Preserves legend)
  print(p72)

  # 2. Interactive output for exploration
  ggplotly(p72, tooltip = "text")
} else {
  cat("No bond angle data available.\n")
}
```

## 8. Coordination Number Analysis

### 8.1 ECon Coordination Number

Continuous CN output directly by the ECon algorithm without additional occupancy weights.

```{r viz-regular-cn, fig.width=10, fig.height=10}
reg_cn_agg <- final_bonds_data[, .(MeanLength = mean(Distance), Count = .N), 
                               by = .(Element1, Z1, CoordinationNumber)]
reg_cn_agg <- reg_cn_agg[Count >= 10]
reg_cn_agg[, Element1 := fct_reorder(Element1, Z1)]

p81 <- ggplot(reg_cn_agg, aes(x = factor(round(CoordinationNumber, 1)), y = Element1,
                              text = paste("Element:", Element1, "<br>ECon CN:", round(CoordinationNumber, 1),
                                           "<br>Avg Dist:", round(MeanLength, 3), "Å",
                                           "<br>Count:", Count))) +
  geom_point(aes(size = Count, color = MeanLength)) +
  scale_color_viridis_c(option = "magma", direction = -1, name = "Mean Length (Å)") +
  scale_size_continuous(range = c(2, 8), breaks = scales::breaks_pretty(n = 5), name = "Count (Observations)") +
  labs(title = "Bond Length vs ECon CN", x = "ECon CN (Rounded for Grouping)", y = "Central Element") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# 1. Static
print(p81)

# 2. Interactive
ggplotly(p81, tooltip = "text")
```

### 8.2 Weighted CN

This view re-analyzes the data by adjusting for crystallographic site occupancies. High Geometric environments often condense down into standard structural motifs.

```{r viz-weighted-cn-z, fig.width=10, fig.height=10}
wcn_agg <- final_bonds_data[, .(MeanLength = mean(Distance), Count = .N), 
                            by = .(Element1, Z1, RoundedWeightedCN)]
wcn_agg <- wcn_agg[Count >= 10 & RoundedWeightedCN > 0]
wcn_agg[, Element1 := fct_reorder(Element1, Z1)]

p82 <- ggplot(wcn_agg, aes(x = factor(RoundedWeightedCN), y = Element1,
                           text = paste("Element:", Element1, "<br>W-CN:", RoundedWeightedCN,
                                        "<br>Avg Dist:", round(MeanLength, 3), "Å",
                                        "<br>Count:", Count))) +
  geom_point(aes(size = Count, color = MeanLength)) +
  scale_color_viridis_c(option = "magma", direction = -1, name = "Mean Length (Å)") +
  scale_size_continuous(range = c(2, 8), breaks = scales::breaks_pretty(n = 5), name = "Count (Observations)") +
  labs(title = "Bond Length vs Weighted CN", x = "Weighted CN", y = "Central Element") +
  theme_minimal()

# 1. Static
print(p82)

# 2. Interactive
ggplotly(p82, tooltip = "text")
```

### 8.3 Weighted CN (Sorted by Global Bond Length)

Re-ordering the Y-axis by the overall median bond length of the element helps rapidly identify which elements form the smallest/tightest coordination environments versus large, highly coordinated cation sites.

```{r viz-weighted-cn-len, fig.width=10, fig.height=10}
wcn_agg_len <- copy(wcn_agg)
elem_means <- wcn_agg_len[, .(GlobalMean = mean(MeanLength)), by = Element1]
wcn_agg_len <- merge(wcn_agg_len, elem_means, by="Element1")
wcn_agg_len[, Element1 := fct_reorder(Element1, GlobalMean)]

p83 <- ggplot(wcn_agg_len, aes(x = factor(RoundedWeightedCN), y = Element1,
                               text = paste("Element:", Element1, "<br>W-CN:", RoundedWeightedCN,
                                            "<br>Avg Dist:", round(MeanLength, 3), "Å"))) +
  geom_point(aes(size = Count, color = MeanLength)) +
  scale_color_viridis_c(option = "magma", direction = -1, name = "Mean Length (Å)") +
  scale_size_continuous(range = c(2, 10), breaks = scales::breaks_pretty(n = 5), name = "Count (Observations)") +
  labs(title = "Length vs W-CN (Elements Sorted by Avg Global Length)", x = "Weighted CN", y = "Central Element") +
  theme_minimal()

# 1. Static
print(p83)

# 2. Interactive
ggplotly(p83, tooltip = "text")
```

## 9. Chalcogen-Specific Analysis

Filtering the entire dataset to place the Chalcogens (Sulfur, Selenium, Tellurium) strictly at the center of the geometric queries. In this section, we review Raw Bond Lengths, Normalized Bond Lengths, ECon Coordination Numbers, Weighted CNs, and Bond Angles.

```{r chalcogen-setup}
targets <- c("S", "Se", "Te")
df_center_chalc <- final_bonds_data[Element1 %in% targets]
```

### 9.1 S, Se, Te: Bond Lengths

```{r chalc-center-len, fig.width=15, fig.height=5}
agg_center <- df_center_chalc[, .(MeanDist = mean(Distance), N=.N), by=.(Element1, Z1, Element2, Z2)]
agg_center <- agg_center[N >= 3]
agg_center[, Element1 := fct_reorder(Element1, Z1)]
agg_center[, Element2 := fct_reorder(Element2, Z2)]

p91 <- ggplot(agg_center, aes(x = Element2, y = Element1,
                              text = paste("Center:", Element1, "<br>Neighbor:", Element2,
                                           "<br>Dist:", round(MeanDist, 3), "Å",
                                           "<br>Obs:", N))) +
  geom_point(aes(size = N, color = MeanDist)) +
  scale_color_viridis_c(option = "magma", direction = -1, name = "Dist (Å)") +
  scale_size_continuous(range = c(3, 10), breaks = scales::breaks_pretty(n = 5), name = "Count (Observations)") +
  labs(title = "Bond Lengths: S, Se, Te Centers", x = "Neighbor Element", y = "Central Chalcogen") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

# 1. Static
print(p91)

# 2. Interactive
ggplotly(p91, tooltip = "text")

# 3. Heatmap
ggplot(agg_center, aes(x = Element2, y = Element1, fill = MeanDist)) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(option = "magma", direction = -1, name = "Dist (Å)") +
  labs(title = "Bond Lengths: S, Se, Te Centers (Heatmap)", 
       subtitle = "Sorted by Atomic Number",
       x = "Neighbor Element", y = "Central Chalcogen") +
  theme_minimal() +
  coord_fixed()
```

### 9.2 S, Se, Te: Normalized Bond Lengths (Distance / Sum of Radii)

```{r chalc-center-norm-len, fig.width=15, fig.height=5}
agg_center_norm <- df_center_chalc[!is.na(NormalizedDistance), .(MeanNormDist = mean(NormalizedDistance), N=.N), by=.(Element1, Z1, Element2, Z2)]
agg_center_norm <- agg_center_norm[N >= 3]
agg_center_norm[, Element1 := fct_reorder(Element1, Z1)]
agg_center_norm[, Element2 := fct_reorder(Element2, Z2)]

p92 <- ggplot(agg_center_norm, aes(x = Element2, y = Element1,
                              text = paste("Center:", Element1, "<br>Neighbor:", Element2,
                                           "<br>Norm Dist:", round(MeanNormDist, 3),
                                           "<br>Obs:", N))) +
  geom_point(aes(size = N, color = MeanNormDist)) +
  scale_color_viridis_c(option = "magma", direction = -1, name = "Norm Dist") +
  scale_size_continuous(range = c(3, 10), breaks = scales::breaks_pretty(n = 5), name = "Count (Observations)") +
  labs(title = "Normalized Bond Lengths (Distance / Sum of Radii): S, Se, Te Centers", x = "Neighbor Element", y = "Central Chalcogen") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

# 1. Static
print(p92)

# 2. Interactive
ggplotly(p92, tooltip = "text")

# 3. Heatmap
ggplot(agg_center_norm, aes(x = Element2, y = Element1, fill = MeanNormDist)) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(option = "magma", direction = -1, name = "Norm Dist") +
  labs(title = "Normalized Bond Lengths: S, Se, Te Centers (Heatmap)", 
       subtitle = "Distance divided by sum of covalent radii",
       x = "Neighbor Element", y = "Central Chalcogen") +
  theme_minimal() +
  coord_fixed()
```

### 9.3 S, Se, Te: ECon Coordination Number

```{r chalc-center-cn, fig.width=15, fig.height=5}
agg_center_cn <- df_center_chalc[, .(MeanCN = mean(CoordinationNumber), N=.N), 
                                 by=.(Element1, Z1, Element2, Z2)]
agg_center_cn <- agg_center_cn[N >= 3]
agg_center_cn[, Element1 := fct_reorder(Element1, Z1)]
agg_center_cn[, Element2 := fct_reorder(Element2, Z2)]

p93 <- ggplot(agg_center_cn, aes(x = Element2, y = Element1,
                                 text = paste("Center:", Element1, "<br>Neighbor:", Element2,
                                              "<br>Avg ECon CN:", round(MeanCN, 2),
                                              "<br>Obs:", N))) +
  geom_point(aes(size = N, color = MeanCN)) +
  scale_color_viridis_c(option = "magma", direction = -1, name = "Avg CN") +
  scale_size_continuous(range = c(3, 10), breaks = scales::breaks_pretty(n = 5), name = "Count (Observations)") +
  labs(title = "ECon CN: S, Se, Te Centers", x = "Neighbor Element", y = "Central Chalcogen") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

# 1. Static
print(p93)

# 2. Interactive
ggplotly(p93, tooltip = "text")

# 3. Heatmap
ggplot(agg_center_cn, aes(x = Element2, y = Element1, fill = MeanCN)) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(option = "magma", direction = -1, name = "Avg CN") +
  labs(title = "ECon Coordination Number: S, Se, Te Centers", 
       subtitle = "Sorted by Atomic Number",
       x = "Neighbor Element", y = "Central Chalcogen") +
  theme_minimal() +
  coord_fixed()
```

### 9.4 S, Se, Te: Weighted CN (Special Aggregation)

In this visualization, we map **Specific WCN** (the fraction of the coordination environment contributed strictly by the specific element pairing) and **Total WCN** (the complete occupancy-corrected size of the environment).
* **Color:** The average specific element-element Weighted CN.
* **Size:** The average Total Weighted CN for the site.

```{r chalc-center-wcn, fig.width=15, fig.height=5}
# 1. Extract the actual Total WCN per central site
atom_totals <- final_bonds_data[, .(
  Total_WCN = first(CN_Weighted)
), by = .(file_name, Atom1, Element1, Z1)]

# 2. Extract Exact Specific WCN (Summing the true occupancies of the bonded neighbors)
atom_specifics <- final_bonds_data[, .(
  Specific_WCN = sum(Occupancy)
), by = .(file_name, Atom1, Element1, Z1, Element2, Z2)]

# 3. Merge Specific WCN with Total WCN for plotting
df_plot_wcn <- merge(atom_specifics, atom_totals, by = c("file_name", "Atom1", "Element1", "Z1"))

# 4. Filter down to strictly Chalcogen centers
targets <- c("S", "Se", "Te")
df_plot_chalc <- df_plot_wcn[Element1 %in% targets]

# Reorder factors by Atomic Number for the plot axes
df_plot_chalc[, Element1 := fct_reorder(Element1, Z1)]
df_plot_chalc[, Element2 := fct_reorder(Element2, Z2)]

# --- CRITICAL FIXES FOR SIZING ---
# Sort data so LARGEST bubbles are plotted FIRST (background), 
# and SMALLEST bubbles are plotted LAST (foreground)
df_plot_chalc <- df_plot_chalc[order(-Total_WCN)]

# Extract maximums to anchor the legends properly
max_tot <- ceiling(max(df_plot_chalc$Total_WCN, na.rm = TRUE))
max_spec <- ceiling(max(df_plot_chalc$Specific_WCN, na.rm = TRUE))

p94 <- ggplot(df_plot_chalc, aes(x = Element2, y = Element1,
                                 text = paste("File:", file_name,
                                              "<br>Center:", Element1, "(Site:", Atom1, ")",
                                              "<br>Neighbor:", Element2,
                                              "<br>Specific WCN (Color):", round(Specific_WCN, 2),
                                              "<br>Total Env Size (Size):", round(Total_WCN, 2)))) +
  # Apply vertical jitter (offset)
  geom_jitter(aes(size = Total_WCN, color = Specific_WCN), 
              width = 0.15, height = 0.25, alpha = 0.7) +
  scale_color_viridis_c(option = "magma", direction = -1, 
                        name = "Specific WCN",
                        limits = c(0, max_spec)) +
  # DRASTICALLY WIDEN THE SIZE RANGE so differences are visually obvious
  scale_size_continuous(range = c(1, 7), 
                        name = "Total WCN Env Size",
                        breaks = seq(2, max_tot, by = 2),
                        limits = c(0, max_tot)) +
  labs(title = "Weighted CN Analysis: Specific vs Total Environment Size", 
       subtitle = "Individual Coordination Environments (Largest Values Plotted in Background)",
       x = "Neighbor Element", y = "Central Chalcogen") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

# 1. Static
print(p94)

# 2. Interactive
ggplotly(p94, tooltip = "text")

# 3. Aggregate Heatmap 
agg_center_wcn <- df_plot_chalc[, .(MeanSpecificWCN = mean(Specific_WCN), N=.N), 
                                  by=.(Element1, Z1, Element2, Z2)]
agg_center_wcn <- agg_center_wcn[N >= 3]

ggplot(agg_center_wcn, aes(x = Element2, y = Element1, fill = MeanSpecificWCN)) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(option = "magma", direction = -1, name = "Mean Specific WCN") +
  labs(title = "Mean Specific Weighted CN: S, Se, Te Centers", 
       subtitle = "Sorted by Atomic Number (Disorder Corrected)",
       x = "Neighbor Element", y = "Central Chalcogen") +
  theme_minimal() +
  coord_fixed()
```

To eliminate the vertical overlap here is a split analysis, with the above chart separated into three distinct plots for **Sulfur**, **Selenium**, and **Tellurium**. 

### 9.4.1 Specific WCN: Sulfur (S) Centers

```{r chalc-wcn-s, fig.width=15, fig.height=5}
# Filter for S only
df_s <- df_plot_chalc[Element1 == "S"]

p_s <- ggplot(df_s, aes(x = Element2, y = Element1,
                        text = paste("File:", file_name,
                                     "<br>Site:", Atom1,
                                     "<br>Neighbor:", Element2,
                                     "<br>Specific WCN:", round(Specific_WCN, 2),
                                     "<br>Total Site WCN:", round(Total_WCN, 2)))) +
  geom_jitter(aes(size = Total_WCN, color = Specific_WCN), 
              width = 0.2, height = 0.5, alpha = 0.6) +
  scale_color_viridis_c(option = "magma", direction = -1, name = "Specific WCN", limits = c(0, max_spec)) +
  scale_size_continuous(range = c(1, 8), name = "Total WCN", limits = c(1, max_tot)) +
  labs(title = "Sulfur Coordination Environments", x = "Neighbor Element", y = "Central S") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

print(p_s)
ggplotly(p_s, tooltip = "text")
```

### 9.4.2 Specific WCN: Selenium (Se) Centers

```{r chalc-wcn-se, fig.width=15, fig.height=5}
# Filter for Se only
df_se <- df_plot_chalc[Element1 == "Se"]

p_se <- ggplot(df_se, aes(x = Element2, y = Element1,
                          text = paste("File:", file_name,
                                       "<br>Site:", Atom1,
                                       "<br>Neighbor:", Element2,
                                       "<br>Specific WCN:", round(Specific_WCN, 2),
                                       "<br>Total Site WCN:", round(Total_WCN, 2)))) +
  geom_jitter(aes(size = Total_WCN, color = Specific_WCN), 
              width = 0.2, height = 0.5, alpha = 0.6) +
  scale_color_viridis_c(option = "magma", direction = -1, name = "Specific WCN", limits = c(0, max_spec)) +
  scale_size_continuous(range = c(1, 8), name = "Total WCN", limits = c(1, max_tot)) +
  labs(title = "Selenium Coordination Environments", x = "Neighbor Element", y = "Central Se") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

print(p_se)
ggplotly(p_se, tooltip = "text")
```

### 9.4.3 Specific WCN: Tellurium (Te) Centers

```{r chalc-wcn-te, fig.width=15, fig.height=5}
# Filter for Te only
df_te <- df_plot_chalc[Element1 == "Te"]

p_te <- ggplot(df_te, aes(x = Element2, y = Element1,
                          text = paste("File:", file_name,
                                       "<br>Site:", Atom1,
                                       "<br>Neighbor:", Element2,
                                       "<br>Specific WCN:", round(Specific_WCN, 2),
                                       "<br>Total Site WCN:", round(Total_WCN, 2)))) +
  geom_jitter(aes(size = Total_WCN, color = Specific_WCN), 
              width = 0.2, height = 0.5, alpha = 0.6) +
  scale_color_viridis_c(option = "magma", direction = -1, name = "Specific WCN", limits = c(0, max_spec)) +
  scale_size_continuous(range = c(1, 8), name = "Total WCN", limits = c(1, max_tot)) +
  labs(title = "Tellurium Coordination Environments", x = "Neighbor Element", y = "Central Te") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

print(p_te)
ggplotly(p_te, tooltip = "text")
```

### 9.5 S, Se, Te: Bond Angles

```{r chalc-center-angles, fig.width=15, fig.height=5}
if (exists("all_angles_data") && nrow(all_angles_data) > 0) {
  df_angles_sub <- all_angles_data[Element1 %in% targets]
  
  agg_angles_chalc <- df_angles_sub[, .(MeanAngle = mean(Angle), N=.N), by=.(Element1, Z1, Element2, Z2)]
  agg_angles_chalc <- agg_angles_chalc[N >= 3]
  agg_angles_chalc[, Element1 := fct_reorder(Element1, Z1)]
  agg_angles_chalc[, Element2 := fct_reorder(Element2, Z2)]
  
  p95 <- ggplot(agg_angles_chalc, aes(x = Element2, y = Element1,
                                 text = paste("Center:", Element1, "<br>Neighbor:", Element2,
                                              "<br>Avg Angle:", round(MeanAngle, 1), "°",
                                              "<br>Obs:", N))) +
    geom_point(aes(size = N, color = MeanAngle)) +
    scale_color_viridis_c(option = "magma", name = "Avg Angle (°)") +
    scale_size_continuous(range = c(3, 10), breaks = scales::breaks_pretty(n = 5), name = "Count (Observations)") +
    labs(title = "Bond Angles: S, Se, Te Centers", x = "Neighbor Element", y = "Central Chalcogen") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
    
  # 1. Static
  print(p95)
  
  # 2. Interactive
  ggplotly(p95, tooltip = "text")
  
  # 3. Heatmap
  ggplot(agg_angles_chalc, aes(x = Element2, y = Element1, fill = MeanAngle)) +
    geom_tile(color = "white") +
    scale_fill_viridis_c(option = "magma", name = "Avg Angle (°)") +
    labs(title = "Bond Angles: S, Se, Te Centers (Heatmap)", 
         x = "Neighbor Element", y = "Central Chalcogen") +
    theme_minimal() +
    coord_fixed()
} else {
  cat("No angle data available.\n")
}
```

## 10. Distribution Analysis

This visualization compares the distribution of bond lengths for the most common coordination numbers (e.g., 4 and 6), allowing us to see the “spread” of bond distances for specific elements. Because ECon generates continuous decimal values, we round them to the nearest integer strictly for visualization grouping.

```{r viz-distributions, fig.width=12, fig.height=6}
# --- Distribution Analysis (Violin Plot) ---

# 1. Round continuous ECon CNs and filter for common Coordination Numbers
violin_data <- copy(final_bonds_data)
violin_data[, RoundedCN := round(CoordinationNumber)]
common_cns <- c(4, 6) # Focusing on Tetrahedral and Octahedral equivalent environments
violin_data <- violin_data[RoundedCN %in% common_cns]

# 2. Filter to top 20 most frequent elements to keep the x-axis clean
top_elements <- violin_data[, .N, by = Element1][order(-N)][1:20, Element1]
violin_data <- violin_data[Element1 %in% top_elements]

# 3. Plot
ggplot(violin_data, aes(x = reorder(Element1, Distance), y = Distance)) +
  geom_violin(aes(fill = factor(RoundedCN)), scale = "width", alpha = 0.7, position = position_dodge(width = 0.8)) +
  geom_boxplot(aes(group = interaction(Element1, RoundedCN)), 
               width = 0.1, outlier.shape = NA, fill="white", position = position_dodge(width = 0.8)) +
  scale_fill_brewer(palette = "Dark2", name = "Rounded ECon CN") +
  labs(title = "Distribution of Bond Lengths",
       subtitle = "Comparing Coordination Numbers ~4 and ~6 for Top Elements",
       x = "Element (Sorted by Median Distance)",
       y = "Bond Distance (Å)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## 11. Advanced: Length vs Angle (Grouped by Neighbor Family)

Faceted by S, Se, Te. Neighbors are colored by **standard IUPAC periodic table groups** (Alkali metals, Reactive nonmetals, Lanthanides, etc.), while preserving the split between **Early TMs** and **Late TMs**.

*(Note: Data extraction steps below use `copy()` extensively to guarantee safety and isolation from the global objects, preventing any interactive environment bleeding).*

```{r viz-advanced-scatter, fig.width=12, fig.height=6, warning=FALSE, message=FALSE}
# --- 1. Expanded Group Metadata ---

extended_groups <- data.table(
  Symbol = c(
    # Alkali metals
    "Li", "Na", "K", "Rb", "Cs", "Fr",
    
    # Alkaline earth metals
    "Be", "Mg", "Ca", "Sr", "Ba", "Ra",
    
    # Early TMs (Now includes up to Group 7: Mn, Tc, Re)
    "Sc", "Y", "Ti", "Zr", "Hf", "V", "Nb", "Ta", "Cr", "Mo", "W", "Mn", "Tc", "Re",
    
    # Late TMs (Groups 8-12)
    "Fe", "Ru", "Os", "Co", "Rh", "Ir", "Ni", "Pd", "Pt", "Cu", "Ag", "Au", "Zn", "Cd", "Hg",
    
    # Post-transition metals
    "Al", "Ga", "In", "Tl", "Sn", "Pb", "Bi", "Po",
    
    # Metalloids
    "B", "Si", "Ge", "As", "Sb", "Te", "At",
    
    # Reactive nonmetals
    "H", "C", "N", "P", "O", "S", "Se", "F", "Cl", "Br", "I",
    
    # Noble gases
    "He", "Ne", "Ar", "Kr", "Xe", "Rn",
    
    # Lanthanides
    "La", "Ce", "Pr", "Nd", "Pm", "Sm", "Eu", "Gd", "Tb", "Dy", "Ho", "Er", "Tm", "Yb", "Lu",
    
    # Actinides
    "Ac", "Th", "Pa", "U", "Np", "Pu", "Am", "Cm", "Bk", "Cf", "Es", "Fm", "Md", "No", "Lr",
    
    # Unknown properties
    "Rf", "Db", "Sg", "Bh", "Hs", "Mt", "Ds", "Rg", "Cn", "Nh", "Fl", "Mc", "Lv", "Ts", "Og"
  ),
  GroupLabel = c(
    rep("Alkali metals", 6),
    rep("Alkaline earth metals", 6),
    rep("Early TMs", 14),
    rep("Late TMs", 15),
    rep("Post-transition metals", 8),
    rep("Metalloids", 7),
    rep("Reactive nonmetals", 11),
    rep("Noble gases", 6),
    rep("Lanthanides", 15),
    rep("Actinides", 15),
    rep("Unknown properties", 15)
  )
)

if (exists("all_angles_data") && nrow(all_angles_data) > 0) {
  # A. Filter Angles for S, Se, Te centers ONLY
  target_centers <- c("S", "Se", "Te")
  df_angles_sub <- copy(all_angles_data[CentralAtom %like% "^(S|Se|Te)([0-9]|$|_)"])
  
  # Extract Base Symbols
  df_angles_sub[, CenterSym := str_extract(CentralAtom, "^[A-Za-z]+")]
  df_angles_sub[, NeighborSym1 := str_extract(Neighbor1, "^[A-Za-z]+")]
  df_angles_sub[, NeighborSym2 := str_extract(Neighbor2, "^[A-Za-z]+")]
  
  # Strict filter to keep only S, Se, Te
  df_angles_sub <- df_angles_sub[CenterSym %in% target_centers]
  
  # B. Prepare Raw ECon Bonds List for Distance Fetching
  # We use final_bonds_data, which contains the raw lengths of all ECon bonds.
  # This merge does NOT filter the angles, it merely acts as a lookup table to get lengths.
  raw_econ_bonds <- copy(unique(final_bonds_data[, .(file_name, Atom1, Atom2, Distance)], by = c("file_name", "Atom1", "Atom2")))
  
  # C. Fetch Dist1 (Central to Neighbor1)
  plot_data <- merge(
    df_angles_sub,
    raw_econ_bonds,
    by.x = c("file_name", "CentralAtom", "Neighbor1"),
    by.y = c("file_name", "Atom1", "Atom2"),
    allow.cartesian = TRUE
  )
  setnames(plot_data, "Distance", "Dist1")
  
  # D. Fetch Dist2 (Central to Neighbor2)
  plot_data <- merge(
    plot_data,
    raw_econ_bonds,
    by.x = c("file_name", "CentralAtom", "Neighbor2"),
    by.y = c("file_name", "Atom1", "Atom2")
  )
  setnames(plot_data, "Distance", "Dist2")
  
  # E. Calculate Raw Distance between Neighbor1 and Neighbor2 (Law of Cosines)
  plot_data[, Angle_rad := Angle * pi / 180]
  plot_data[, Dist_N1_N2 := sqrt(Dist1^2 + Dist2^2 - 2 * Dist1 * Dist2 * cos(Angle_rad))]
  
  # F. Merge Covalent Radii to determine the physical threshold
  # Merge for Neighbor1
  plot_data <- merge(
    plot_data,
    custom_radii_values[, .(Symbol, Radius)],
    by.x = "NeighborSym1",
    by.y = "Symbol",
    all.x = TRUE
  )
  setnames(plot_data, "Radius", "Radius_N1")
  
  # Merge for Neighbor2
  plot_data <- merge(
    plot_data,
    custom_radii_values[, .(Symbol, Radius)],
    by.x = "NeighborSym2",
    by.y = "Symbol",
    all.x = TRUE
  )
  setnames(plot_data, "Radius", "Radius_N2")
  
  # G. Apply Physical Distance Check (The 15% Ghost Filter)
  # Removes pairs where the distance between the two neighbors is less than the
  # sum of their covalent radii multiplied by (1 - 0.15)
  plot_data <- plot_data[!is.na(Radius_N1) & !is.na(Radius_N2)]
  plot_data[, expected_dist := Radius_N1 + Radius_N2]
  plot_data[, lower_bound := expected_dist * (1 - 0.15)]
  plot_data <- plot_data[Dist_N1_N2 >= lower_bound]
  
  # H. Merge Neighbor Group Info (Categorized by Neighbor 1)
  plot_data <- merge(
    plot_data,
    extended_groups,
    by.x = "NeighborSym1",
    by.y = "Symbol",
    all.x = TRUE
  )
  
  # Remove NAs (elements not in our group list)
  plot_data <- plot_data[!is.na(GroupLabel)]
  
  # I. Sort Central Atom for Plotting Order
  plot_data[, CenterSym := factor(CenterSym, levels = c("S", "Se", "Te"))]
  
  # --- 3. Plotting ---
  p11 <- ggplot(plot_data, aes(
    x = Dist1,
    y = Angle,
    text = paste(
      "File:",
      file_name,
      "Center:",
      CentralAtom,
      "<br>Neighbor1:",
      Neighbor1,
      "<br>Neighbor2:",
      Neighbor2,
      "<br>Group:",
      GroupLabel,
      "<br>N1-N2 Dist:",
      round(Dist_N1_N2, 3),
      "Å",
      "<br>Allowed Minimum:",
      round(lower_bound, 3),
      "Å",
      "<br>Angle:",
      Angle,
      "°"
    )
  )) +
    geom_point(aes(color = GroupLabel), alpha = 0.4, size = 1) +
    facet_wrap(~ CenterSym, scales = "free", ncol = 3) +
    scale_color_viridis_d(option = "turbo", name = "Neighbor Family") +
    labs(
      title = "Bond Length vs Angle Correlation",
      subtitle = "Filtered for Occupancy Artifacts (Physical Neighbor Distance Check) | Faceted by Central Atom",
      x = "Bond Length (Å)",
      y = "Bond Angle (°)"
    ) +
    theme_minimal() +
    theme(legend.position = "bottom") +
    guides(color = guide_legend(override.aes = list(alpha = 1, size = 3)))
  
  # 1. Static
  print(p11)
  
  # 2. Interactive (toWebGL helps performance with 10k+ points)
  toWebGL(ggplotly(p11, tooltip = "text")) %>%
    layout(legend = list(
      orientation = "h",
      x = 0,
      y = -0.2
    ))
  
} else {
  cat("No angle data available to plot.\n")
}
```

## 12. Multi-Algorithm Analysis on Chalcogenide Subset

This section samples ~50 chalcogenide CIF files from our previously filtered list and runs them through a multi-algorithm coordination analysis. While the entire notebook was rebuilt around **ECon**, here we benchmark it directly alongside the other main bonding algorithms provided by `crystract`. We extract both geometric/standard CNs and Weighted CNs.

```{r multi-algo-test, message=FALSE, warning=FALSE}
# --- 1. Load required libraries ---
library(crystract)
library(data.table)

set_radii_data(custom_radii_values)

# --- 2. Sample ~50 Chalcogenide Files ---
# Set seed for reproducibility
set.seed(123)

# Destination for the test files
dest_dir <- "/Users/donngo/Documents/test"
if (!dir.exists(dest_dir)) {
  dir.create(dest_dir, recursive = TRUE)
} else {
  # Clean directory to avoid mixing old test runs
  invisible(file.remove(list.files(dest_dir, full.names = TRUE)))
}

# Use `target_files` (generated in Section 3) to guarantee chemistry logic
num_to_sample <- min(50, length(target_files))
selected_files <- sample(target_files, num_to_sample)

# Copy selected files to destination folder explicitly 
dest_files <- file.path(dest_dir, basename(selected_files))
file.copy(from = selected_files, to = dest_files, overwrite = TRUE)

# --- 3. Run the batch analysis ---
cat("\nStarting multi-algorithm batch analysis on CIF files...\n")
results <- analyze_cif_files(
  file_paths = dest_dir,
  tolerance = 1e-4,
  bonding_algorithms = c("minimum_distance", "brunner", "econ", "voronoi", "crystal_nn"),
  calculate_bond_angles = FALSE,       
  perform_error_propagation = FALSE,   
  workers = 9                          
)

# --- 4. Extract and Combine Geometric Coordination Numbers ---
output_dir <- "/Users/donngo/Documents/cif_results"
if (!dir.exists(output_dir)) dir.create(output_dir, recursive = TRUE)

# Find all data tables that start with "cn_" (i.e. coordination number results)
cn_columns <- grep("^cn_", names(results), value = TRUE)

cat("\nExtracting Geometric Coordination Numbers and combining into a single CSV...\n")

# Initialize the combined table with the file names
combined_dt <- data.table(file = results$file_name)

# Loop over each algorithm column
for (col in cn_columns) {
  combined_dt[[col]] <- sapply(results[[col]], function(dt) {
    if (is.data.frame(dt) && nrow(dt) > 0 && "CoordinationNumber" %in% names(dt)) {
      # Sort alphabetically by Atom so the sequence matches across algorithms
      if ("Atom" %in% names(dt)) dt <- dt[order(Atom)]
      
      # Round output to handle decimals provided by continuous algorithms like ECon
      cn_rounded <- round(dt$CoordinationNumber, 3)
      return(paste(cn_rounded, collapse = ", "))
    } else {
      return(NA_character_)
    }
  })
}

# Save the final combined CN table
combined_file_path <- file.path(output_dir, "combined_cn_crystract.csv")
fwrite(combined_dt, combined_file_path)
cat(sprintf("Saved multi-algorithm CN table with %d rows to: %s\n", nrow(combined_dt), basename(combined_file_path)))


# --- 5. Extract and Combine Weighted Coordination Numbers (WCN) ---
cat("\nExtracting Weighted Coordination Numbers (WCN) and combining into a single CSV...\n")

# Initialize the combined table for WCN with the file names
combined_wcn_dt <- data.table(file = results$file_name)

# Loop over each algorithm column
for (col in cn_columns) {
  combined_wcn_dt[[col]] <- sapply(results[[col]], function(dt) {
    if (is.data.frame(dt) && nrow(dt) > 0 && "WeightedCoordinationNumber" %in% names(dt)) {
      # Sort alphabetically by Atom to maintain identical ordering to the geometric CN list
      if ("Atom" %in% names(dt)) dt <- dt[order(Atom)]
      
      # Round to 3 decimal places to keep the comma-separated strings readable
      wcn_rounded <- round(dt$WeightedCoordinationNumber, 3)
      
      return(paste(wcn_rounded, collapse = ", "))
    } else {
      return(NA_character_)
    }
  })
}

# Save the final combined WCN table
combined_wcn_file_path <- file.path(output_dir, "combined_wcn_crystract.csv")
fwrite(combined_wcn_dt, combined_wcn_file_path)
cat(sprintf("Saved WCN table with %d rows to: %s\n", nrow(combined_wcn_dt), basename(combined_wcn_file_path)))


cat("\nAll operations finished successfully!\n")
```


## 13. High Coordination Number Diagnostics (WCN > 12)

To address the request to double-check structural validity, this section isolates the specific `.cif` files that yielded a Weighted Coordination Number (WCN) strictly greater than 12. It displays the sites and automatically copies the underlying `.cif` files to an export directory for easy sharing.

```{r extract-high-wcn-files, message=FALSE, warning=FALSE}
if (exists("final_bonds_data") && nrow(final_bonds_data) > 0) {
  
  # Filter for sites with Weighted CN > 12
  high_wcn_data <- unique(final_bonds_data[CN_Weighted > 12, .(
    File = file_name,
    Site = Atom1,
    Element = Element1,
    `Geometric CN` = round(CoordinationNumber, 3),
    `Weighted CN` = round(CN_Weighted, 3)
  )])
  
  # Sort by Weighted_CN descending to see the most extreme cases first
  high_wcn_data <- high_wcn_data[order(-`Weighted CN`)]
  
  unique_high_wcn_files <- unique(high_wcn_data$File)
  
  cat(sprintf("Found %d atomic sites across %d unique files where WCN > 12.\n", 
              nrow(high_wcn_data), length(unique_high_wcn_files)))
  
  # 1. Display the table interactively for the HTML report
  if(nrow(high_wcn_data) > 0) {
    p_high_wcn <- datatable(
      high_wcn_data, 
      extensions = 'Buttons',
      options = list(
        pageLength = 10, 
        scrollX = TRUE,
        dom = 'Bfrtip',
        buttons = c('csv', 'excel')
      ),
      caption = "Sites with Weighted Coordination Number > 12"
    )
    print(p_high_wcn)
    
    # 2. Export the physical CIF files for review
    # The baseline cif_directory is set in Section 3
    export_dir <- "/Users/donngo/Documents/high_wcn_cifs"
    if(!dir.exists(export_dir)) dir.create(export_dir, recursive = TRUE)
    
    files_to_copy <- file.path(cif_directory, unique_high_wcn_files)
    existing_files <- files_to_copy[file.exists(files_to_copy)]
    
    if(length(existing_files) > 0) {
      file.copy(from = existing_files, to = export_dir, overwrite = TRUE)
      cat(sprintf("\nSuccessfully copied %d CIF files to %s.", 
                  length(existing_files), export_dir))
    } else {
      cat("\nWarning: Could not locate the original .cif files to copy. Check 'cif_directory' path.\n")
    }
  }
  
} else {
  cat("No bond data available to extract high WCN files.")
}
```
