---
title: "Advanced Workflow: High-Throughput Analysis of Clathrate Structures"
author: "Don Ngo, Julia Maria Hubner, Anirudh Prabhu"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
vignette: >
  %\VignetteIndexEntry{Advanced Workflow: High-Throughput Analysis of Clathrate Structures}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
# This chunk sets up the environment for the analysis.
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, warning = FALSE,
  fig.width = 10, fig.height = 7, fig.align = 'center'
)

# Load all necessary libraries
library(crystract)
library(data.table)
library(ggplot2)
library(plotly)
library(DT)
library(htmlwidgets)
library(knitr)

# Load the package's built-in covalent radii data
data(covalent_radii, package = "crystract")
```

## 1. Introduction and Objective

This vignette demonstrates an advanced, robust workflow using the `crystract` package to analyze a large, complex dataset. While the introductory vignette focuses on the step-by-step analysis of a single, well-ordered crystal, this tutorial tackles the challenges of real-world, high-throughput analysis, specifically for disordered structures like clathrates.

The goal is to process a dataset of clathrate structures to calculate a representative "weighted average network bond distance" for each one and explore its relationship with the lattice parameter 'a'.

### The Robust Pipeline for Disordered Structures

Clathrate structures often feature site-occupancy disorder and guest atoms within a host framework. A naive distance calculation can lead to non-physical "ghost" distances. To handle this, we will implement a specific, multi-step filtering pipeline for each crystal:

1.  **Filter Guests:** Isolate the host framework by removing guest atoms.
2.  **Filter Ghosts:** Remove physically implausible distances using `filter_ghost_distances()`.
3.  **Identify Bonds:** Determine the true bonded pairs from the cleaned data using `minimum_distance()`.
4.  **Calculate Average:** Compute the final `calculate_weighted_average_network_distance()` only on the confirmed, physical bonds.

This **Guests -> Ghosts -> Bonds -> Average** sequence ensures the final metric is accurate and physically meaningful.

## 2. Preparing the Data

### 2.1. Obtaining and Organizing ICSD Files

Due to licensing restrictions, the dataset of clathrate CIF files from the Inorganic Crystal Structure Database (ICSD) is **not** bundled with this package. To reproduce this workflow, you must download the data via your institutional license and organize it specifically for this script.

#### Step 1: Download Required Files
A list of the specific ICSD Collection Codes required for this analysis (and their corresponding structural categories) can be found in `required_ICSD_files.csv`, located in the root of the `ml-crystals` repository. Download the `.cif` file for each entry.

#### Step 2: Create the Directory Structure
The analysis script relies on the folder names to determine the **structural category** (which determines the Wyckoff sites used for calculations). You must create a main directory (e.g., `my_clathrate_data`) and create subdirectories named exactly after the categories found in the CSV file.

The structure must look like this:

```text
/my_clathrate_data/              <-- Main Directory
│
├── 6c-16i-24k/                  <-- Subdirectory (Category Name)
│     ├── ICSD_CollCode12345.cif <-- CIF files belonging to this category
│     ├── ICSD_CollCode12346.cif
│     └── ...
└── ...
```

**Note:** The subdirectory names (e.g., `6c-16i-24k`) are parsed by the script to determine which atoms belong to the host network. Ensure these folder names match the CSV categories exactly.

### 2.2. Loading the Data

In the chunk below, you must update the `cif_root_dir` variable to point to the location where you organized the files.

```{r load-files, echo=TRUE, eval=FALSE}
# IMPORTANT: Update this path to point to your created main data directory
cif_root_dir <- "C:/path/to/my_clathrate_data"

# Check if the directory exists
if (!dir.exists(cif_root_dir)) {
  stop("CRITICAL ERROR: Could not find the data directory. Please update 'cif_root_dir'.")
}

# Identify category subdirectories
category_dirs <- list.dirs(path = cif_root_dir, full.names = TRUE, recursive = FALSE)

# Map files to their category based on the folder name
file_map_list <- lapply(category_dirs, function(dir) {
  full_paths <- list.files(path = dir, pattern = "\\.cif$", full.names = TRUE, ignore.case = TRUE)
  if (length(full_paths) == 0) return(NULL)
  
  # The folder name becomes the category identifier
  data.table(
    file_path = full_paths, 
    file_name = basename(full_paths), 
    category = basename(dir)
  )
})

file_map <- rbindlist(file_map_list, fill = TRUE)
all_cif_paths <- if (!is.null(file_map) && nrow(file_map) > 0) file_map$file_path else character(0)

cat("Found", length(all_cif_paths), "total CIF files across", length(category_dirs), "categories.\n")

if (length(all_cif_paths) == 0) {
    stop("Execution stopped: No CIF files were found inside the specified directory.")
}
```

## 3. Optimized Batch Analysis and Custom Pipeline

### 3.1. Step 1: Optimized Raw Data Extraction

We first run `analyze_cif_files()`. For efficiency and to enable our custom pipeline, we disable the built-in bonding calculations (`bonding_algorithms = "none"`). This performs the most intensive step—calculating all interatomic distances—just once.

```{r run-batch-analysis-optimized, echo=TRUE, eval=FALSE}
# Run the extraction and distance calculations
analysis_results <- analyze_cif_files(
  file_paths = all_cif_paths,
  perform_extraction = TRUE,
  perform_calcs_and_transforms = TRUE,
  bonding_algorithms = "none",        # Skip standard bonding logic
  calculate_bond_angles = FALSE,      # Skip angles for speed
  perform_error_propagation = FALSE
)

cat("Optimized analysis complete. Raw results table has", nrow(analysis_results), "rows.\n")
```

### 3.2. Step 2: Applying the "Guests -> Ghosts -> Bonds -> Average" Pipeline

Now, we apply our core filtering logic by looping through the results for each file. This section extracts the Wyckoff targets from the `category` column (which came from your folder names) and applies the specific logic required for that clathrate type.

```{r process-data, echo=TRUE, eval=FALSE}
# --- Timing Start ---
timing_processing <- system.time({
  
  # Define atoms to exclude (Guests)
  guest_atoms <- c("Na", "K", "Rb", "Cs", "Sr", "Ba", "Eu")
  all_removed_ghosts <- list()
  
  process_file_results <- function(i) {
    current_filename <- analysis_results$file_name[i]
    
    # Retrieve the category established in the loading step
    category_info <- file_map[file_name == current_filename]
    
    if (is.null(current_filename) || nrow(category_info) == 0) return(NULL)
    
    distances <- analysis_results$distances[[i]]
    coords <- analysis_results$atomic_coordinates[[i]]
    
    if (is.null(distances) || is.null(coords) || nrow(distances) == 0) return(NULL)
    
    # --- Pipeline Steps ---
    category <- category_info$category
    
    # Parse the folder name (e.g., "6c-16i-24k") to get target Wyckoff sites
    # This logic assumes folder names are formatted as "Site1-Site2-Site3"
    target_wyckoff_symbols <- strsplit(gsub("\\+M_on_.*", "", category), "-")[[1]]
    
    # 1. Filter Guest Atoms
    distances <- filter_by_elements(distances, coords, guest_atoms)
    if (nrow(distances) == 0) return(NULL)
    
    # 2. Filter Ghost Distances
    ghost_filter_result <- filter_ghost_distances(distances, coords, margin = 0.1)
    cleaned_distances <- ghost_filter_result$kept
    
    removed_table <- ghost_filter_result$removed
    if (nrow(removed_table) > 0) {
      removed_table[, file := current_filename]
      all_removed_ghosts[[length(all_removed_ghosts) + 1]] <<- removed_table
    }
    if (nrow(cleaned_distances) == 0) return(NULL)
    
    # 3. Identify Bonds
    bonded_pairs <- minimum_distance(cleaned_distances, delta = 0.1)
    if (nrow(bonded_pairs) == 0) return(NULL)
    
    # 4. Calculate Weighted Average
    weighted_avg <- calculate_weighted_average_network_distance(bonded_pairs, coords, target_wyckoff_symbols)
    if (is.na(weighted_avg)) return(NULL)
    
    return(
      data.table(
        file = current_filename,
        category = category,
        lattice_parameter_a = analysis_results$unit_cell_metrics[[i]]$`_cell_length_a`,
        weighted_distance = weighted_avg
      )
    )
  }
  
  plot_data_list <- lapply(1:nrow(analysis_results), process_file_results)
  plot_data <- rbindlist(plot_data_list, use.names = TRUE, fill = TRUE)
  plot_data <- na.omit(plot_data)
  removed_ghosts_summary <- rbindlist(all_removed_ghosts, fill = TRUE)
}) # --- Timing End ---

cat("Data processing complete.\n")
cat("Final plot table has", nrow(plot_data), "entries.\n")
if (exists("removed_ghosts_summary") && nrow(removed_ghosts_summary) > 0) {
  cat(
    nrow(removed_ghosts_summary),
    "non-physical 'ghost' distances were identified and removed.\n"
  )
}
```

## 4. Quality Control: Review of Removed "Ghost" Distances

A key advantage of this workflow is the ability to inspect the distances that `filter_ghost_distances()` removed. This is crucial for validating the filtering logic and understanding the nature of disorder in the dataset.

```{r show-removed-distances, echo=TRUE, eval=FALSE}
if (exists("removed_ghosts_summary") && nrow(removed_ghosts_summary) > 0) {
  sample_size <- min(1000, nrow(removed_ghosts_summary))
  cat(paste("Displaying a sample of the first", sample_size, "removed distances below:\n"))
  
  datatable(
    removed_ghosts_summary[1:sample_size, ],
    caption = paste("Sample of Removed Ghost Distances (", sample_size, " of ", nrow(removed_ghosts_summary), " total)"),
    rownames = FALSE, extensions = 'Buttons',
    options = list(pageLength = 10, dom = 'Bfrtip', buttons = c('copy', 'csv')),
    colnames = c("Atom 1", "Atom 2", "Removed Distance (Å)", "Expected (Å)", "Lower Bound (Å)", "Upper Bound (Å)", "Reason", "File")
  )
} else {
  cat("No 'ghost' distances were detected during the analysis.")
}
```

## 5. Visualizing and Exploring Final Results

After applying the robust pipeline, we can now visualize the cleaned data.

### 5.1. Interactive Plot

```{r generate-plot, echo=TRUE, eval=FALSE}
# Only generate the plot if the data processing was successful
if (exists("plot_data") && nrow(plot_data) > 0) {
  p <- ggplot(plot_data, aes(x = lattice_parameter_a, y = weighted_distance, color = category, shape = category, text = file)) +
    geom_point(alpha = 0.65, size = 2.5) +
    geom_smooth(aes(group = 1), method = "lm", se = FALSE, color = "black", linetype = "dotted", fullrange = TRUE) +
    labs(
      title = "Average Network Bond Length vs. Lattice Parameter 'a'",
      subtitle = "Data points are colored by structural category",
      x = "Lattice Parameter a (Å)",
      y = "Weighted Average Network Bond Length (Å)",
      color = "Structural Category", shape = "Structural Category"
    ) +
    theme_bw(base_size = 14) +
    theme(legend.position = "bottom", legend.title = element_text(face = "bold"))

  interactive_plot <- ggplotly(p, tooltip = c("x", "y", "text", "color"))
  interactive_plot
} else {
  cat("No data available to generate plot.")
}
```

### 5.2. Interactive Data Table

The final, cleaned data used for the plot is provided below in a searchable and exportable table.

```{r show-final-data-table, echo=TRUE, eval=FALSE}
if (exists("plot_data") && nrow(plot_data) > 0) {
  datatable(
    plot_data,
    caption = "Final Processed Data for Clathrate Structures",
    rownames = FALSE, filter = 'top', extensions = 'Buttons',
    options = list(pageLength = 15, dom = 'Bfrtip', buttons = c('copy', 'csv', 'excel')),
    colnames = c("File", "Category", "Lattice 'a' (Å)", "Weighted Dist (Å)")
  )
} else {
  cat("No final data to display.")
}
```

## 6. Conclusion

This vignette has demonstrated how to leverage the `crystract` package to build a powerful, customized pipeline for high-throughput analysis. By organizing your own ICSD data into a structured directory tree, you can utilize the package's robust processing capabilities. The "Guests -> Ghosts -> Bonds -> Average" pattern presented here ensures that the final summary metrics are physically meaningful, even when dealing with real-world disordered structures.
